### 2008-08-19 doc/manual.tex
docs
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@618 ce7aba0d-9f81-40a6-a495-cecae84b0515

\documentclass[11pt]{article}
\usepackage{fancyhdr}
\usepackage[dvips]{graphicx} 
\usepackage{amsmath,amssymb} 
\usepackage{epsfig}
\usepackage{calc}

\newcommand{\simname}{BookSim~}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Setup the margin sizes.

\evensidemargin = 0in
\oddsidemargin = 0in
\textwidth = 6.5in

\topmargin = -0.5in
\textheight = 9in

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Ted Jiang, Brian Towles and William J. Dally}
\title{\simname 2.0 User's Guide}

\begin{document}

\maketitle
\tableofcontents

\pagestyle{fancy}
%\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\fancyhf{} % delete current setting for header and footer
\fancyhead[LE,RO]{\bfseries\thepage}
\fancyhead[LO]{\bfseries\rightmark}
\fancyhead[RE]{\bfseries\leftmark}
\renewcommand{\headrulewidth}{0.5pt}
\renewcommand{\footrulewidth}{0.5pt}
\addtolength{\headheight}{0.5pt} % make space for the rule
\cfoot{\small\today}
\fancypagestyle{plain}{%
    \fancyhf{} % get rid of headers on plain pages
    \renewcommand{\headrulewidth}{0pt} % and the line
    \renewcommand{\footrulewidth}{0pt} % and the line
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newenvironment{opt_list}[1]{\begin{list}{}{\renewcommand{\makelabel}[1]%
{\texttt{##1}\hfil}\settowidth{\labelwidth}{\texttt{#1}}\setlength{\leftmargin}%
{\labelwidth+\labelsep}}}{\end{list}}

\section{Introduction}

This document describes the use of the \simname interconnection
network simulator.  The simulator is designed as a companion to the
textbook ``Principles and Practices of Interconnection Networks''
(PPIN) published by Morgan Kaufmann (ISBN: 0122007514) and it is
assumed that is reader is familiar with the material covered in that
text.

This user guide is fairly brief as, with most simulators, the best way
to learn and {\it understand} the simulator is to study the code.
Most of the simulator's components are designed to be modular so tasks
such as adding a new routing algorithm, topology, or router
microarchitecture should not require a complete redesign of the code.
Once you have downloaded the code, compiled it, and run a simple
example (Section~\ref{sec:get_started}), the more detailed examples of
Section~\ref{sec:examples} give a good overview of the capabilities of
the simulator.  A list of configuration options is provided in
Section~\ref{sec:config_params} for reference.

\section{Getting started}
\label{sec:get_started}

\subsection{Downloading and building the simulator}
\label{sec:download}

The latest version of the simulator is available from
\texttt{http://cva.stanford.edu} as a compressed tar archive.  UNIX/Linux
users can extract this archive using the tar utility
\begin{verbatim}
  tar xvfz booksim-1.0.tar.gz  
\end{verbatim}
Windows users can use a compression program such as WinZip to extract
the archive.

The simulator itself is written in C++ and has been specifically
tested with GNU's G++ compiler (version $\ge3$).  In addition, both a
LEX and YACC tool (also known as FLEX and BISON) are needed to create
the configuration parser.  These are standard tools in any UNIX/Linux
development environment.  It is suggested that Windows users download
the CYGWIN versions (\texttt{http://www.cygwin.com}) of these UNIX
development tools to simplify their compilation process.  The
\texttt{Makefile} should be edited so that the first lines give the
paths to the tools.  At Stanford, for example, the compiler, YACC, and
LEX are stored in the \texttt{/usr/pubsw/bin} directory.  The
\texttt{Makefile} reflects this:
\begin{verbatim}
CPP    = /usr/pubsw/bin/g++
YACC   = /usr/pubsw/bin/byacc -d
LEX    = /usr/pubsw/bin/flex
\end{verbatim}
Then, the simulator can be compiled by running \texttt{make} in the
directory that contains the \texttt{Makefile}.

\subsection{Running a simulation}
\label{sec:run_example}

The syntax of the simulator is simply
\begin{verbatim}
  booksim [configfile]
\end{verbatim}
The optional parameter \texttt{configfile} is a file that contains
configuration information for the simulator.  So, for example, to
simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube)
network on uniform traffic, a configuration such as the one shown in
Figure~\ref{fig:config_example} could be used.  This particular
configuration is stored in \texttt{examples/torus88}.

\begin{figure}
\begin{verbatim}
  // Topology
  topology = torus;
  k        = 8;
  n        = 2;

  // Routing
  routing_function = dim_order;

  // Flow control
  num_vcs = 2;

  // Traffic
  traffic        = uniform;
  injection_rate = 0.15;
\end{verbatim}
\caption{Example configuration file for simulating a 8-ary 2-cube
network.}
\label{fig:config_example}
\end{figure}

In addition to specifying the topology, the configuration file also
contains basic information about the routing algorithm, flow control,
and traffic.  This simple example uses dimension-order routing and, to
ensure deadlock-freedom of this routing function in the torus, two
virtual channels are required.  The \texttt{injection\_rate} parameter
is added to tell the simulator to inject 0.15 flits per simulation
cycle per node.  Because the simulator operates at the flit level,
most parameters are specified in units of flits as is the case with
the \texttt{injection\_rate}.  Also, any line of the configuration
that begins with \texttt{//} is treated as a comment and ignored by
the simulator.  A detailed list of configuration parameters is given in
Section~\ref{sec:config_params}.

\subsection{Simulation output}

Continuing our example, running the torus simulation produces the
output shown in Figure~\ref{fig:sim_output}.  Each simulation has
three basic phases: warm up, measurement, and drain.  The length of
the warm up and measurement phases is a multiple of a basic sample
period (defined by \texttt{sample\_period} in the configuration).  As
shown in the figure, the current latency and throughput (rate of
accepted packets) for the simulation is printed after each sample
period.  The overall throughput is determined by the lowest throughput
of all the destination in the network, but the average throughput is
also displayed.

\begin{figure}
\begin{verbatim}
%=================================
% Average latency = 6.02008
% Accepted packets = 0.11 at node 52 (avg = 0.147094)
% latency change    = 1
% throughput change = 1

...

% Warmed up ...
%=================================
% Average latency = 6.0796
% Accepted packets = 0.119 at node 5 (avg = 0.148266)
% latency change    = 0.00562457
% throughput change = 0.00379387

...

% Draining all recorded packets ...
% Draining remaining packets ...
====== Traffic class 0 ======
Overall average latency = 6.09083 (1 samples)
Overall average accepted rate = 0.149475 (1 samples)
Overall min accepted rate = 0.138551 (1 samples)
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

After the warm up periods have passed, the simulator prints the
``\texttt{Warmed up}'' message and resets all the simulation statistics.
Then, the measurement phase begins and statistics continue to be
reported after each sample period.  Once the measurement periods have
passed, all the measurement packets are drained from the network
before final latency and throughput numbers are reported.  Details of
the configuration parameters used to control the length of the
simulation phases are covered in Section~\ref{sec:sim_params}.

\section{Examples}
\label{sec:examples}

One of the most basic performance measures of any interconnection
network is its latency versus offered load.
Figure~\ref{fig:lat_vs_load} shows a simple configuration file for
making this measurement in a 8-ary 2-mesh network under the transpose
traffic pattern.  This configuration was used to generate Figure 25.2
in PPIN.  The particular configuration accounts for some small delays
and pipelining of the input-queued router and also introduces a small
input speedup to account for any inefficiencies in allocation.  By
running simulations for many increments of \texttt{injection\_rate},
the average latency curve can be found.  Then, to compare the
performance of dimension-order routing against several other routing
algorithms, for example, the \texttt{routing\_function} option can be
changed.

\begin{figure}
\begin{verbatim}
// Topology

topology = mesh;
k = 8;
n = 2;

// Routing

routing_function = dim_order;

// Flow control

num_vcs     = 8;
vc_buf_size = 8;

wait_for_tail_credit = 1;

// Router architecture

vc_allocator = islip;
sw_allocator = islip;
alloc_iters  = 1;

credit_delay   = 2;
routing_delay  = 1;
vc_alloc_delay = 1;

input_speedup     = 2;
output_speedup    = 1;
internal_speedup  = 1.0;

// Traffic

traffic                = transpose;
const_flits_per_packet = 20;

// Simulation

sim_type       = latency;
injection_rate = 0.1;
\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

Figure~\ref{fig:fly_dist} shows a configuration file that can be used
to determine the distribution of packet latencies in a 2-ary 6-fly
network that uses age-based arbitration.  Note the use of the
\texttt{priority} configuration parameter along with the
\texttt{select} allocators that account for packet priorities.  The
simulator does not output latency distributions by default, but by
editing \texttt{trafficmanager.cpp}, setting the configuration
variable \texttt{DISPLAY\_LAT\_DIST} to true, and recompiling, the
distribution will be displayed at the end of the simulation.  This
technique was used to produced the distribution shown in Figure 25.12
of PPIN.

\begin{figure}
\begin{verbatim}
// Topology

topology = fly;
k = 2;
n = 6;

// Routing

routing_function = dest_tag;

// Flow control

num_vcs     = 8;
vc_buf_size = 8;

wait_for_tail_credit = 1;

// Router architecture

vc_allocator = select;
sw_allocator = select;
alloc_iters  = 1;

credit_delay   = 2;
routing_delay  = 1;
vc_alloc_delay = 1;

input_speedup     = 2;
output_speedup    = 1;
internal_speedup  = 1.0;

// Traffic

traffic                = uniform;
const_flits_per_packet = 20;
priority               = age;

// Simulation

sim_type       = latency;
injection_rate = 0.1;
\end{verbatim}
\caption{A configuration file (\texttt{examples/fly26\_age}) for
finding the distribution of packet latencies using age-based
arbitration.}
\label{fig:fly_dist}
\end{figure}

As a final example, Figure~\ref{fig:single} shows the use of the
special single-node topology to test the performance of a switch
allocator --- in this case, the iSLIP allocator.  The
\texttt{in\_ports} and \texttt{out\_ports} options set up a simulation
of an $8\times 8$ crossbar.

\begin{figure}
\begin{verbatim}
// Topology

topology  = single;
in_ports  = 8;
out_ports = 8;

// Routing

routing_function = single;

// Flow control

vc_allocator = islip;
sw_allocator = islip;
alloc_iters  = 2;

num_vcs     = 8;
vc_buf_size = 1000;

wait_for_tail_credit = 0;

// Simulation

sim_type       = latency;
injection_rate = 0.1;
\end{verbatim}
\caption{A single-node configuration file (\texttt{examples/single})
for testing the performance of a switch allocator.}
\label{fig:single}
\end{figure}

\section{Configuration parameters}
\label{sec:config_params}

All information used to configure a simulation is passed through a
configuration file as illustrated by the example in
Section~\ref{sec:run_example}.  This section lists the existing
configuration parameters --- a user can incorporate additional options
by changing the \texttt{booksim\_config.cpp} file.

\subsection{Topologies}
\label{sec:topos}

The \texttt{topology} parameter determines the underlying topology of the
network and the simulator supports four basic topologies:
\begin{opt_list}{single}
\item[fly] A $k$-ary $n$-fly (butterfly) topology. The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter
determines the network's dimension.

\item[mesh] A $k$-ary $n$-mesh (mesh) topology. The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter determines
the network's dimension.

\item[single] A network with a single node, used for testing single
router performance.  The number of input and output ports for the node
is determined by the \texttt{in\_ports} and \texttt{out\_ports} parameters,
respectively.

\item[torus] A $k$-ary $n$-cube (torus) topology.  The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter determines
the network's dimension.
\end{opt_list}

Both the \texttt{mesh} and \texttt{torus} topologies support the
addition of random link failures with the \texttt{link\_failures}
parameter.  The value of \texttt{link\_failures} determines the number
of channels that are randomly removed from the topology and are thus
no longer available for forwarding packets.  Moreover, the
randomization for failed channels is controlled by selecting an
integer value for the \texttt{fail\_seed} parameter --- a fixed seed
gives a fixed set of failed channels, independent of other
randomization in the simulation.  Also, note that only certain routing
functions support this feature (see Section~\ref{sec:routing_algs}).

\subsection{Routing algorithms}
\label{sec:routing_algs}

The \texttt{routing\_function} parameter selects a routing algorithm
for the topology.  Many routing algorithms need multiple virtual
channels for deadlock freedom (VCDF).

\begin{opt_list}{dim\_order\_bal}

\item[dim\_order] Dimension-order routing.  Works for the
\texttt{mesh} topology (1 VCDF) and for the \texttt{torus} topology (2
VCDF).

\item[dim\_order\_bal] Dimension-order routing for the
\texttt{torus} topology with a more balanced use of VCs to
avoid deadlock (2 VCDF).

\item[dim\_order\_ni] A non-interfering version of
dimension-order routing.  Works on the \texttt{torus} or \texttt{mesh}
topology and requires one VC per network terminal.

\item[min\_adapt] A minimal adaptive routing algorithm for
the \texttt{mesh} topology (2 VCDF) and for the \texttt{torus}
topology (3 VCDF).

\item[planar\_adapt] Planar-adaptive routing for the
\texttt{mesh} topology (2 VCDF).  Supports routing around failed channels.

\item[romm] ROMM routing for the \texttt{mesh} (2 VCDF).
Load is balanced by routing in two phases: one from the source to a
random intermediate node in the minimal quadrant and a second from the
intermediate to the destination.

\item[romm\_ni] A non-interfering version of ROMM routing for
the \texttt{mesh} that requires one VC per network terminal.

\item[single] A dummy routing function used for the
\texttt{single} topology.

\item[valiant] Valiant's randomized routing algorithm for the
\texttt{mesh} (2 VCDF) and \texttt{torus} (4 VCDF) topology.

\item[valiant\_ni] A non-interfering version of Valiant's algorithm
for the \texttt{torus} that requires 4 VCs per network terminal.

\end{opt_list}

Also, the simulator code is structured so that additional routing
algorithms can be added with minimal changes to the overall simulator
(see the \texttt{routefunc.cpp} file in the simulator's source code).

\subsection{Flow control}

The simulator supports basic virtual-channel flow control with
credit-based backpressure.

\begin{opt_list}{wait\_for\_tail\_credit}

\item[num\_vcs] The number of virtual channels per physical channel.

\item[vc\_buf\_size] The depth of each virtual in flits.

\item[voq] If non-zero, use virtual-output queuing.  With virtual
output queuing, a separate virtual channel is assigned to each
destination in the network.  This option is most useful when used with
a non-interfering routing algorithm (Section~\ref{sec:routing_algs}).

\item[wait\_for\_tail\_credit] If non-zero, do not reallocate a virtual
channel until the tail flit has left that virtual channel.  This
conservative approach prevents a dependency from being formed between
two packets sharing the same virtual channel in succession.
\end{opt_list}

\subsection{Router organizations}

The simulator also supports two different router microarchitectures.
The input-queued router follows the general organization described in
PPIN while the event-driven router is modeled after the router used in
the Avici TSR and described in U.S. Patent 6,370,145.  The
microarchitecture is selected using the \texttt{router} option.  Also,
both routers share a small set of options.

\begin{opt_list}{internal\_speedup}
\item[credit\_delay] The processing delay (in cycles) for a credit.
Does not include the wire delay for transmitting the credit.

\item[internal\_speedup] An arbitrary speedup of the internals of the
routers over the channel transmission rate.  For example, a speedup
1.5 means that, on average, 1.5 flits can be forwarded by the router
in the time required for a single flit to be transmitted across a
channel.  Also, the configuration parser expects a floating point
number for this field, so integer speedups should also include a
decimal point (e.g. ``2.0'').

\item[output\_delay] The processing delay incurred in the output queue
of a router.
\end{opt_list}

\subsubsection{The input-queued router}
\label{sec:iq_router}

The input-queued router (\texttt{router = iq}) follows the pipeline
described in PPIN of route computation, virtual-channel allocation,
switch allocation, and switch traversal.  There are several options
specific to the input-queued router.

\begin{opt_list}{st\_prepare\_delay}

\item[input\_speedup] An integer speedup of the input ports in space.
A speedup of 2, for example, gives each input two input ports into the
crossbar.  Access to these ports is statically allocated based on the
virtual channel number: virtual channel $v$ at input $i$ is connected
to port $i \cdot s + (v \mod s)$ for an input speedup of $s$.

\item[output\_speedup] An integer speedup of the output ports in
space.  Similar to \texttt{input\_speedup}

\item[routing\_delay] The delay (in cycles) of route computation.

\item[sw\_allocator] The type of allocator used for switch allocation.
See Section~\ref{sec:alloc} for a list of the possible allocators.

\item[sw\_alloc\_delay] The delay (in cycles) of switch allocation.

\item[vc\_allocator] The type of allocator used for virtual-channel
allocation.  See Section~\ref{sec:alloc} for a list of the possible
allocators.

\item[vc\_alloc\_delay] The delay (in cycles) of virtual-channel
allocation.

\end{opt_list}

\subsubsection{The event-driven router}
\label{sec:event_router}

The event-driven router (\texttt{router = event}) is a
microarchitecture designed specifically to support a large number of
virtual channels (VCs) efficiently.  Instead of continuously polling
the state of the virtual channels, as in the input-queued router, only
changes in VC state are tracked.  The efficiency then comes from the
fact that the number of state changes per cycle is constant and
independent of the number of VCs.

\subsection{Allocators}
\label{sec:alloc}

Many of the allocators used in the simulator are configurable (see
the input-queued router in Section~\ref{sec:iq_router}) and several
allocation algorithms are available.
\begin{opt_list}{wavefront}

\item[max\_size] Maximum-size matching. 
\item[islip] iSLIP separable allocator.
\item[pim] Parallel iterative matching separable allocator.
\item[loa] Lonely output allocator.
\item[wavefront] Wavefront matching.
\item[select] Priority-based allocator.  Allocation is performed as in
iSLIP, but with preference towards higher priority packets (see
\texttt{priority} option in Section~\ref{sec:traffic}).

\end{opt_list}

Allocation can also be improved by performing multiple iterations of
the algorithm and the number of iterations is controlled by the
\texttt{alloc\_iters} parameter.

\subsection{Traffic}
\label{sec:traffic}

The rate at which flits are injected into the simulator is set using
the \texttt{injection\_rate} option.  The simulator's cycle time is a
flit cycle, the time it takes a single flit to be injected at a
source, and the injection rate is specified in flits per flit cycle.
For example, setting \texttt{injection\_rate = 0.25} means that each
source injects a new flit one of every four simulator cycles.  The
injection process can also be specified as either Bernoulli
(\texttt{injection\_process = bernoulli}) or an on-off process
(\texttt{injection\_process = on\_off}).  The burstiness of the latter
injection process is controlled via the \texttt{burst\_alpha} and
\texttt{burst\_beta} parameter.  See PPIN Section 24.2.2 for a
description of the on-off process and its parameters.

The unit of injection is packets, which may be comprised of many
flits.  The number of flits per packet is set using the
\texttt{const\_flits\_per\_packet} option.  Each packet may also have an
associated priority, either age-based (\texttt{age}) or none
(\texttt{none}), as specified by the \texttt{priority} option.

The simulator also supports several different traffic patterns that
are specified using the \texttt{traffic} option.  To describe these
patterns, we use the same notation of PPIN Section 3.2: $s_i$ ($d_i$)
denotes the $i^\textrm{th}$ bit of the source (destination) address
whereas $s_x$ ($d_x$) denotes the $x^\textrm{th}$ radix-$k$ digit of
the source (destination) address.  The bit length of an address is $b
= \log_2 N$, where $N$ is the number of nodes in the network.

\begin{opt_list}{transpose}
\item[uniform] Each source sends an equal amount of traffic to each
destination (\texttt{traffic = uniform}).
\item[bitcomp] Bit complement. $d_i = \neg s_i$.
\item[bitrev] Bit reverse. $d_i = s_{b-i-1}$.
\item[shuffle] $d_i = s_{i-1 \mod b}$.
\item[transpose] $d_i = s_{i+b/2 \mod b}$.
\item[tornado] $d_x = s_x + \lceil k/2 \rceil - 1 \mod k$.
\item[neighbor] $d_x = s_x + 1 \mod k$.
\item[randperm] Random permutation.  A fixed permutation traffic
pattern is chosen uniformly at random from the set of all
permutations.  The seed used to generate this permutation is set by
the \texttt{perm\_seed} option.  So, randomly selecting values for
\texttt{perm\_seed} gives a random sampling of permutation while a
fixed value of \texttt{perm\_seed} allows the same permutation to be
used for several experiments.
\end{opt_list}

\subsection{Simulation parameters}
\label{sec:sim_params}

The duration and other aspects of a simulation are controlled using
the set of simulation parameters.

\begin{opt_list}{warmup\_periods}

\item[sim\_type] A simulation can either focus on
\texttt{throughput} or \texttt{latency}.  The key difference between
these two types is that a \texttt{latency} simulation will wait for
all measurement packets to drain before ending the simulation to
ensure an accurate latency measurement.  In \texttt{throughput}
simulations, this final drain step is eliminated to allow simulation
of networks operating beyond their saturation point.

\item[sample\_period] The sample period is expressed in simulator
cycles and is used as a multiplier when specifying the warm-up length
of a simulation and the maximum number of samples.  Also, intermediate
statistics are displayed once every \texttt{sample\_period} cycles.

\item[warmup\_periods] The length of the simulator warm up expressed
as a multiple of the \texttt{sample\_period}.  After warming up, all
statistics counters are reset.

\item[max\_samples] The total length of simulation expressed as a
multiple of the \texttt{sample\_period}.

\item[latency\_thres] If the sampled latency of the current simulation
exceeds \texttt{latency\_thres}, the simulation is immediately ended.

\item[sim\_count] The number of back-to-back simulations to run for the
given configuration.  Useful for creating ensemble averages of
particular statistics.

\item[seed] A random seed for the simulation.

\item[reorder] A non-zero value indicates that packet order should be
maintained and reordering time is accounted for in the overall latency.

\end{opt_list}

\appendix

\section{Random number generation}

The simulator uses Knuth's integer and floating point pseudorandom
number generators.  These algorithms and their explanations appear in
``The Art of Computer Programming: Seminumerical Algorithms''.

\end{document}

### 2008-08-28 doc/manual.tex
Added physical subnetworks. Added progressive adaptive in UGAL. Non-minimal and minimal traffic classes in valiant and UGAL FBFly. Deadlock warning counter. Flip a coin for generation always. Note: It appears to deadlock if wait_for_tail_flits is disabled.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@638 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 
-\author{Ted Jiang, Brian Towles and William J. Dally}
+\author{Ted Jiang, Brian Towles, George Michelogiannakis and William J. Dally}
 \title{\simname 2.0 User's Guide}
@@
 
+\subsection{Physical sub-networks}
+\label{sec:physical_subnets}
+
+The \texttt{physical\_subnetworks} parameter defines the number of physical sub-networks present in the network (defaults to one).
+All sub-networks receive the same configuration parameters and thus are identical. Traffic sources
+maintain an injection queue for each sub-network. The packet generation process is unaffected. It enqueues generated packets into the
+proper sub-network queue according to a division function in the traffic manager. At every cycle, flits at the head of each queue attempt to be injected.
+Traffic destinations can eject one flit from each sub-network each cycle.
+
 \subsection{Routing algorithms}

### 2008-08-28 networks/flatfly_onchip.hpp

@@ -38,6 +38,8 @@ public:
   static void RegisterRoutingFunctions() ;
   double Capacity( ) const;
   void InsertRandomFaults( const Configuration &config );
+
+  static short half_vcs;
 };
 
 

### 2008-10-03 doc/manual.tex
Fixed a memory corruption in the separable allocator. Still does not function correctly though.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@734 ce7aba0d-9f81-40a6-a495-cecae84b0515
restore unix line endings
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@737 ce7aba0d-9f81-40a6-a495-cecae84b0515
minor updates
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@738 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 
-\author{Ted Jiang, Brian Towles, George Michelogiannakis and William J. Dally}
+\author{Ted Jiang, George Michelogiannakis, Daniel Becker, Brian Towles and William J. Dally}
 \title{\simname 2.0 User's Guide}
@@
 
+\subsubsection{The NoC router}
+\label{sec:noc_router}
+
+The NoC router (\texttt{router = noc}) is a simplified variant of the input-queued router.
+It does not support input or output speedup.
+Also, it does not support multi-cycle allocation, but does support output-first allocation.
+
 \subsubsection{The event-driven router}
@@
 \item[loa] Lonely output allocator.
-\item[wavefront] Wavefront matching.
+\item[wavefront] Wavefront allocator.
+\item[matrix] Separable input-first allocator built from matrix arbiters.
+\item[round\_robin] Separable input-first allocator built from round-robin arbiters.
 \item[select] Priority-based allocator.  Allocation is performed as in

### 2008-10-06 doc/manual.tex
docs and examples
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@740 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 \begin{verbatim}
-  tar xvfz booksim-1.0.tar.gz  
+
+  tar xvfz booksim-2.0.tar.gz  
 \end{verbatim}
-Windows users can use a compression program such as WinZip to extract
+
+Windows users can use a compression program such as Winrar to extract
 the archive.
@@
 CPP    = /usr/pubsw/bin/g++
-YACC   = /usr/pubsw/bin/byacc -d
+YACC   = /usr/pubsw/bin/bison -d
 LEX    = /usr/pubsw/bin/flex
@@
 \begin{verbatim}
-  booksim [configfile]
+  ./booksim [configfile]
 \end{verbatim}
@@
 network on uniform traffic, a configuration such as the one shown in
-Figure~\ref{fig:config_example} could be used.  This particular
-configuration is stored in \texttt{examples/torus88}.
+Figure~\ref{fig:config_example} could be used. This particular configuration is stored in examples/torus88.
 
@@
   // Flow control
-  num_vcs = 2;
+  num_vcs = 4;
 
@@
 contains basic information about the routing algorithm, flow control,
-and traffic.  This simple example uses dimension-order routing and, to
-ensure deadlock-freedom of this routing function in the torus, two
-virtual channels are required.  The \texttt{injection\_rate} parameter
-is added to tell the simulator to inject 0.15 flits per simulation
-cycle per node.  Because the simulator operates at the flit level,
-most parameters are specified in units of flits as is the case with
-the \texttt{injection\_rate}.  Also, any line of the configuration
+and traffic.  This simple example uses dimension-order routing and and four
+virtual channels.  The \texttt{injection\_rate} parameter
+is added to tell the simulator to inject 0.15 packets per simulation
+cycle per node.  The packets are default to size of 1 flit.  Also, any line of the configuration
 that begins with \texttt{//} is treated as a comment and ignored by
@@
 \begin{verbatim}
-%=================================
 % Average latency = 6.02008
@@
 % throughput change = 1
+>>>>>>> variant B
+BEGIN Configuration File
+Name: examples/torus88
+// Topology
+topology = torus;
+k = 8;
+n = 2;
+// Routing
+routing_function = dim_order;
+// Flow control
+num_vcs = 4;
+// Traffic
+traffic = uniform;
+injection_rate = 0.15;
+
+END Configuration File
+Unable to open flit watch file, continuing with simulation
+0%=================================
+% Average latency = 10.6252
+% Accepted packets = 0.119 at node 5 (avg = 0.148875)
+lat(1) = 10.6252;
+bins = [ 0 ...999];
+freq = [ 0 0 124 16 497 71 920 166 1336 303 1553 436 1351 
+406 926 328 537 213 173 77 43 24 13 6 5 2 1 1 0 ...]
 
@@
 
-% Warmed up ...
-%=================================
-% Average latency = 6.0796
-% Accepted packets = 0.119 at node 5 (avg = 0.148266)
-% latency change    = 0.00562457
-% throughput change = 0.00379387
+% Warmed up ...Time used is 3000 cycles
 
@@
 % Draining remaining packets ...
+Time taken is 6031 cycles
 ====== Traffic class 0 ======
-Overall average latency = 6.09083 (1 samples)
-Overall average accepted rate = 0.149475 (1 samples)
-Overall min accepted rate = 0.138551 (1 samples)
+Overall average latency = 10.658 (1 samples)
+Overall average accepted rate = 0.148383 (1 samples)
+Overall min accepted rate = 0.134279 (1 samples)
+Average hops = 5.01172 (57148 samples)
 \end{verbatim}
@@
 
-After the warm up periods have passed, the simulator prints the
+After the warm up periods have passed (default to 3$\times$sample\_period), the simulator prints the
 ``\texttt{Warmed up}'' message and resets all the simulation statistics.
@@
 sim_type       = latency;
-injection_rate = 0.1;
+injection_rate = 0.005;
 \end{verbatim}
@@
 \label{fig:lat_vs_load}
-\end{figure}
-
-Figure~\ref{fig:fly_dist} shows a configuration file that can be used
-to determine the distribution of packet latencies in a 2-ary 6-fly
-network that uses age-based arbitration.  Note the use of the
-\texttt{priority} configuration parameter along with the
-\texttt{select} allocators that account for packet priorities.  The
-simulator does not output latency distributions by default, but by
-editing \texttt{trafficmanager.cpp}, setting the configuration
-variable \texttt{DISPLAY\_LAT\_DIST} to true, and recompiling, the
-distribution will be displayed at the end of the simulation.  This
-technique was used to produced the distribution shown in Figure 25.12
-of PPIN.
-
-\begin{figure}
-\begin{verbatim}
-// Topology
-
-topology = fly;
-k = 2;
-n = 6;
-
-// Routing
-
-routing_function = dest_tag;
-
-// Flow control
-
-num_vcs     = 8;
-vc_buf_size = 8;
-
-wait_for_tail_credit = 1;
-
-// Router architecture
-
-vc_allocator = select;
-sw_allocator = select;
-alloc_iters  = 1;
-
-credit_delay   = 2;
-routing_delay  = 1;
-vc_alloc_delay = 1;
-
-input_speedup     = 2;
-output_speedup    = 1;
-internal_speedup  = 1.0;
-
-// Traffic
-
-traffic                = uniform;
-const_flits_per_packet = 20;
-priority               = age;
-
-// Simulation
-
-sim_type       = latency;
-injection_rate = 0.1;
-\end{verbatim}
-\caption{A configuration file (\texttt{examples/fly26\_age}) for
-finding the distribution of packet latencies using age-based
-arbitration.}
-\label{fig:fly_dist}
-\end{figure}
-
-As a final example, Figure~\ref{fig:single} shows the use of the
-special single-node topology to test the performance of a switch
-allocator --- in this case, the iSLIP allocator.  The
-\texttt{in\_ports} and \texttt{out\_ports} options set up a simulation
-of an $8\times 8$ crossbar.
-
-\begin{figure}
-\begin{verbatim}
-// Topology
-
-topology  = single;
-in_ports  = 8;
-out_ports = 8;
-
-// Routing
-
-routing_function = single;
-
-// Flow control
-
-vc_allocator = islip;
-sw_allocator = islip;
-alloc_iters  = 2;
-
-num_vcs     = 8;
-vc_buf_size = 1000;
-
-wait_for_tail_credit = 0;
-
-// Simulation
-
-sim_type       = latency;
-injection_rate = 0.1;
-\end{verbatim}
-\caption{A single-node configuration file (\texttt{examples/single})
-for testing the performance of a switch allocator.}
-\label{fig:single}
 \end{figure}
@@
 configuration parameters --- a user can incorporate additional options
-by changing the \texttt{booksim\_config.cpp} file.
+by changing the \texttt{booksim\_config.cpp} file. Most configuration parameters have default values that should work without user specification
 
@@
 The \texttt{topology} parameter determines the underlying topology of the
-network and the simulator supports four basic topologies:
+network. There is also a set of numerical parameters that describes the size of the networks.
+
+\begin{opt_list}{single}
+\item[k] Network radix, the number of routers per dimension
+\item[n] Network dimension
+\item[c] Network concentration, the number of nodes sharing a single router. Typically set to 1, >1 only in networks that has concentration (i.e. cmesh)
+\item[x] (NoC simulations only) The number of routers in the X dimension. Used to calculate channel latency between routers. 
+\item[y] (NoC simulations only) The number of routers in the y dimension. Used to calculate channel latency between routers. 
+\item[xr] (NoC simulations only) For networks that has c>1, the number of nodes in the x direction per router. Used to calculate channel latency between routers. 
+\item[yr] (NoC simulations only) For networks that has c>1, the number of nodes in the y direction per router. Used to calculate channel latency between routers. 
+\item[limit] Emulate partially filled networks.
+Nodes greater or equal to limit does not generate or receive packets during simulation.
+Though depending on the routing algorithm the routers associated with those nodes may be use to route passing by traffic.
+
+\end{opt_list}
+
+Topologies available in \simname are:
 \begin{opt_list}{single}
@@
 the network's dimension.
+\item[cmesh] Concentrated mesh topology is a A $k$-ary $n$-mesh topology with multiple nodes sharing a single router.
+The \texttt{c} determines the concentration.
+The cmesh topology has the option that turns on "express channels" as described in  by default these channels are turned off.
+
+\item[fat tree] Fat Tree topology with 3 levels.
+Nodes are routers are arranged in a tree structure but the number of links between levels stays constant.
+At the bottom \texttt{k} nodes shares a level 0 router.
+
+\item[flattened butterfly] A topology based on ***
+
+\item[quad tree] A quad tree topology.
+
+\item[tree 4]
+
+\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung.
+
+\item[MECS] A topology based on (***). Currently it only support 1 flit per packet.
+
 \end{opt_list}
@@
 channels for deadlock freedom (VCDF).
-
-\begin{opt_list}{dim\_order\_bal}
-
-\item[dim\_order] Dimension-order routing.  Works for the
-\texttt{mesh} topology (1 VCDF) and for the \texttt{torus} topology (2
-VCDF).
-
-\item[dim\_order\_bal] Dimension-order routing for the
-\texttt{torus} topology with a more balanced use of VCs to
-avoid deadlock (2 VCDF).
-
-\item[dim\_order\_ni] A non-interfering version of
-dimension-order routing.  Works on the \texttt{torus} or \texttt{mesh}
-topology and requires one VC per network terminal.
-
-\item[min\_adapt] A minimal adaptive routing algorithm for
-the \texttt{mesh} topology (2 VCDF) and for the \texttt{torus}
-topology (3 VCDF).
-
-\item[planar\_adapt] Planar-adaptive routing for the
-\texttt{mesh} topology (2 VCDF).  Supports routing around failed channels.
-
-\item[romm] ROMM routing for the \texttt{mesh} (2 VCDF).
-Load is balanced by routing in two phases: one from the source to a
-random intermediate node in the minimal quadrant and a second from the
-intermediate to the destination.
-
-\item[romm\_ni] A non-interfering version of ROMM routing for
-the \texttt{mesh} that requires one VC per network terminal.
-
-\item[single] A dummy routing function used for the
-\texttt{single} topology.
-
-\item[valiant] Valiant's randomized routing algorithm for the
-\texttt{mesh} (2 VCDF) and \texttt{torus} (4 VCDF) topology.
-
-\item[valiant\_ni] A non-interfering version of Valiant's algorithm
-for the \texttt{torus} that requires 4 VCs per network terminal.
-
-\end{opt_list}
-
-Also, the simulator code is structured so that additional routing
-algorithms can be added with minimal changes to the overall simulator
-(see the \texttt{routefunc.cpp} file in the simulator's source code).
+In addition to \texttt{routefunc.cpp}, some topologies source files include additional routing functions.
+Also, the simulator code is structured so that additional routing algorithms can be added with minimal changes to the overall simulator (see the \texttt{routefunc.cpp} file in the simulator's source code).
 
@@
 
-\item[vc\_buf\_size] The depth of each virtual in flits.
+\item[vc\_buf\_size] The depth of each virtual channel in flits.
 
-\item[voq] If non-zero, use virtual-output queuing.  With virtual
-output queuing, a separate virtual channel is assigned to each
-destination in the network.  This option is most useful when used with
-a non-interfering routing algorithm (Section~\ref{sec:routing_algs}).
+%voq is not currently enabled in the simulator !
+%\item[voq] If non-zero, use virtual-output queuing.  With virtual
+%output queuing, a separate virtual channel is assigned to each
+%destination in the network.  This option is most useful when used with
+%a non-interfering routing algorithm (Section~\ref{sec:routing_algs}).
 
@@
 
-The rate at which flits are injected into the simulator is set using
+The two traffic modes, injection rate or batch, can be set by the integer paramter \texttt[use\_read\_write], where 0 means injection mode. Injection mode is on by default.
+
+\subsubsection{Injection mode}
+The rate at which packkets are injected into the simulator is set using
 the \texttt{injection\_rate} option.  The simulator's cycle time is a
 flit cycle, the time it takes a single flit to be injected at a
-source, and the injection rate is specified in flits per flit cycle.
+source, and the injection rate is specified in packets per flit cycle.
 For example, setting \texttt{injection\_rate = 0.25} means that each
-source injects a new flit one of every four simulator cycles.  The
+source injects a new packet one of every four simulator cycles.  The
 injection process can also be specified as either Bernoulli
@@
 The unit of injection is packets, which may be comprised of many
-flits.  The number of flits per packet is set using the
-\texttt{const\_flits\_per\_packet} option.  Each packet may also have an
-associated priority, either age-based (\texttt{age}) or none
-(\texttt{none}), as specified by the \texttt{priority} option.
 
+flits. All packets in injection mode are identical. The number of flits per packet is set using the
+\texttt{const\_flits\_per\_packet} option.
+
+\subsubsection[Batch mode]
+Batch mode is enabled by setting \texttt{use\_read\_write= 1}.
+In batch mode, \texttt[injection\_rate] paramter is ignored.
+Instead, traffic are injected into the network as fast as possible until \texttt{batch\_size} of request packets are generated by each node.
+When a destination node receives a request packet, it returns a reply packet.
+The total number of packets injected in batch mode is 2$\times$ \texttt{batch\_size} per node.
+The parameter \texttt[max\_outstanding] specifies the maximum number of (request or reply) packets waiting at a node to be injected into the network.
+
+\subsubsection[Traffic Patterns]
 The simulator also supports several different traffic patterns that
@@
 of a simulation and the maximum number of samples.  Also, intermediate
-statistics are displayed once every \texttt{sample\_period} cycles.
+statistics are displayed once every \texttt{sample\_period} cycles. This is only applicable in injection mode.
 
@@
 as a multiple of the \texttt{sample\_period}.  After warming up, all
-statistics counters are reset.
+statistics counters are reset. This is only applicable in injection mode.
 
 \item[max\_samples] The total length of simulation expressed as a
-multiple of the \texttt{sample\_period}.
+multiple of the \texttt{sample\_period}. This is only applicable in injection mode.
 
@@
 
-\item[reorder] A non-zero value indicates that packet order should be
-maintained and reordering time is accounted for in the overall latency.
+%This is currently not setup in the traffic manager.
+%\item[reorder] A non-zero value indicates that packet order should be
+%maintained and reordering time is accounted for in the overall latency.
+
+\item[print\_activity] At the end of a simulation using iq\_router, print out the activity for buffer, switch, and channel of the network.
+
+\item[viewer\_trace] The simulator will generate very verbose print out of all activity inside the network.
+This print out should be fed into noc\_viewer for a graphic display of the activity inside the network.
+Currently not working.
+
+\item[watch\_file] Specific flits can have their "watch" status turn on. Require input a file which has flit id listed. 1 id per line.
 

### 2008-10-15 doc/manual.tex
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@744 ce7aba0d-9f81-40a6-a495-cecae84b0515
update allocators
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@763 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 
-\begin{opt_list}{single}
+\begin{opt_list}{topoparams}
+
 \item[k] Network radix, the number of routers per dimension
@@
 \item[y] (NoC simulations only) The number of routers in the y dimension. Used to calculate channel latency between routers. 
-\item[xr] (NoC simulations only) For networks that has c>1, the number of nodes in the x direction per router. Used to calculate channel latency between routers. 
-\item[yr] (NoC simulations only) For networks that has c>1, the number of nodes in the y direction per router. Used to calculate channel latency between routers. 
+\item[xr] (NoC simulations only) For networks that have c>1, the number of nodes in the x direction per router. Used to calculate channel latency between routers. 
+\item[yr] (NoC simulations only) For networks that have c>1, the number of nodes in the y direction per router. Used to calculate channel latency between routers. 
 \item[limit] Emulate partially filled networks.
@@
 
-Topologies available in \simname are:
-\begin{opt_list}{single}
+The channel latency of the network must be configured within the source code of the topology files. All topologies by default have channel latency of 1 cycles. Topologies available in \simname are:
+
+\begin{opt_list}{topologies}
 \item[fly] A $k$-ary $n$-fly (butterfly) topology. The \texttt{k}
@@
 
-\item[MECS] A topology based on (***). Currently it only support 1 flit per packet.
+\item[MECS] A topology based on ***. Currently it only support 1 flit per packet.
 
@@
 
-\begin{opt_list}{wait\_for\_tail\_credit}
+\begin{opt_list}{flowcontrolparams}
 
@@
 
-\begin{opt_list}{internal\_speedup}
+\begin{opt_list}{routerparams}
 \item[credit\_delay] The processing delay (in cycles) for a credit.
@@
 
-\begin{opt_list}{st\_prepare\_delay}
+\begin{opt_list}{iqrouterparams}
 
@@
 
+\item[hold\_switch\_for\_packet]
+
+\item[speculative] Enable speculative switch allocation (i.e., allow switch allocation to occur in parallel with VC allocation for header flits).
+
+\item[filter\_spec\_grants] Determines how speculative grants are masked (\texttt{any\_nonspec\_gnts}: any non-speculative grant inhibits all speculative grants; \texttt{confl\_nonspec\_reqs}: speculative grants are inhibited by conflicting non-speculative requests; \texttt{confl\_nonspec\_gnts}: speculative grants are inhibited by conflicting non-speculative grants).
+
 \item[sw\_allocator] The type of allocator used for switch allocation.
 See Section~\ref{sec:alloc} for a list of the possible allocators.
+
+\item[sw\_alloc\_arb\_type] The type of arbiter to use if a separable input- or output-first allocator is used for switch allocation.
 
@@
 allocators.
+
+\item[vc\_alloc\_arb\_type] The type of arbiter to use if a separable input- or output-first allocator is used for virtual channel allocation.
 
@@
 
-The NoC router (\texttt{router = noc}) is a simplified variant of the input-queued router.
-It does not support input or output speedup.
-Also, it does not support multi-cycle allocation, but does support output-first allocation.
+The NoC router (\texttt{router = noc}) is a simplified variant of the input-queued router. It does not support input or output speedup. Also, it does not support multi-cycle allocation.
 
@@
 allocation algorithms are available.
-\begin{opt_list}{wavefront}
+
+\begin{opt_list}{allocators}
 
@@
 \item[wavefront] Wavefront allocator.
-\item[matrix] Separable input-first allocator built from matrix arbiters.
-\item[round\_robin] Separable input-first allocator built from round-robin arbiters.
+\item[separable\_input\_first] Separable input-first allocator.
+\item[separable\_output\_first] Separable output-first allocator.
 \item[select] Priority-based allocator.  Allocation is performed as in
@@
 
-Allocation can also be improved by performing multiple iterations of
-the algorithm and the number of iterations is controlled by the
-\texttt{alloc\_iters} parameter.
+For the \texttt{islip}, \texttt{pim} and \texttt{select} allocators, allocation can be improved by performing multiple iterations of the algorithm; the number of iterations is controlled by the \texttt{alloc\_iters} parameter.
 
@@
 
-\begin{opt_list}{transpose}
+\begin{opt_list}{trafficpatterns}
 \item[uniform] Each source sends an equal amount of traffic to each
@@
 
-\begin{opt_list}{warmup\_periods}
+\begin{opt_list}{simparams}
 

### 2008-10-15 src/allocators/allocator.hpp

@@ -53,6 +53,7 @@ public:
   static Allocator *NewAllocator( const Configuration &config,
 				  Module *parent, const string& name,
 				  const string &alloc_type, 
+				  const string &arb_type, 
 				  int inputs, int input_speedup,
 				  int outputs, int output_speedup );
 };

### 2008-10-15 src/allocators/separable_output_first.cpp

@@ -1,0 +1,75 @@
+// ----------------------------------------------------------------------
+//
+//  SeparableOutputFirstAllocator: Separable Output-First Allocator
+//
+// ----------------------------------------------------------------------
+
+#include "separable_output_first.hpp"
+
+#include "booksim.hpp"
+#include "arbiter.hpp"
+
+#include <vector>
+#include <iostream>
+#include <string.h>
+
+SeparableOutputFirstAllocator::
+SeparableOutputFirstAllocator( const Configuration& config,
+			       Module* parent, const string& name,
+			       const string& arb_type,
+			       int inputs, int outputs )
+  : SeparableAllocator( config, parent, name, arb_type, inputs, outputs )
+{}
+
+void SeparableOutputFirstAllocator::Allocate() {
+  
+  _ClearMatching() ;
+
+//  cout << "SeparableOutputFirstAllocator::Allocate()" << endl ;
+//  PrintRequests() ;
+  
+  list<sRequest> out_reqs[_outputs];
+  
+  for ( int input = 0 ; input < _inputs ; input++ ) {
+    
+    // Add requests to the output arbiters
+    list<sRequest>::const_iterator it  = _requests[input].begin() ;
+    list<sRequest>::const_iterator end = _requests[input].end() ;
+    while ( it != end ) {
+      const sRequest& req = *it ;
+      if ( req.label > -1 ) {
+	_output_arb[req.port]->AddRequest( input, req.label, req.out_pri );
+      }
+      it++ ;
+    }
+    
+  }
+  
+  for ( int output = 0; output < _outputs; output++ ) {
+    
+    // Execute the output arbiters and propagate the grants to the
+    // input arbiters.
+    int label, pri ;
+    int in = _output_arb[output]->Arbitrate( &label, &pri ) ;
+    
+    if ( in > -1 ) {
+      _input_arb[in]->AddRequest( output, label, pri ) ;
+    }
+  }
+  
+  // Execute the input arbiters.
+  for ( int input = 0 ; input < _inputs ; input++ ) {
+
+    int label, pri ;
+    int output      = _input_arb[input]->Arbitrate( &label, &pri ) ;
+  
+    if ( output > -1 ) {
+      assert( _inmatch[input] == -1 && _outmatch[output] == -1 ) ;
+      _inmatch[input]   = output ;
+      _outmatch[output] = input ;
+      _input_arb[input]->UpdateState() ;
+      _output_arb[output]->UpdateState() ;
+
+    }
+  }
+}

### 2008-11-06 doc/manual.tex
update documentation for new speculation code
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@817 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 
-\item[MECS] A topology based on ***. Currently it only support 1 flit per packet.
+\item[MECS] A topology based on ***. Currently it only supports single-flit packets.
 
@@
 
+\item[alloc\_iters] For the \texttt{islip}, \texttt{pim} and \texttt{select} allocators, allocation can be improved by performing multiple iterations of the algorithm; this parameter controls the number of iterations to be performed for both switch and VC allocation.
+
 \item[sw\_allocator] The type of allocator used for switch allocation.
@@
 
-\item[sw\_alloc\_arb\_type] The type of arbiter to use if a separable input- or output-first allocator is used for switch allocation.
+\item[sw\_alloc\_arb\_type] If the switch allocator is a separable input- or output-first allocator, this parameter selects the type of arbiter to use.
+
+\item[sw\_alloc\_iters] If the switch allocator is of type \texttt{islip}, \texttt{pim} or \texttt{select}, this parameter controls how many allocation iterations should be performed.
+This overrides the global \texttt{alloc\_iters} parameter.
 
@@
 
-\item[vc\_alloc\_arb\_type] The type of arbiter to use if a separable input- or output-first allocator is used for virtual channel allocation.
+\item[vc\_alloc\_arb\_type] If the VC allocator is a separable input- or output-first allocator, this parameter selects the type of arbiter to use.
+
+\item[vc\_alloc\_iters] If the VC allocator is of type \texttt{islip}, \texttt{pim} or \texttt{select}, this parameter controls how many allocation iterations should be performed.
+This overrides the global \texttt{alloc\_iters} parameter.
 
@@
 \end{opt_list}
-
-\subsubsection{The NoC router}
-\label{sec:noc_router}
-
-The NoC router (\texttt{router = noc}) is a simplified variant of the input-queued router. It does not support input or output speedup. Also, it does not support multi-cycle allocation.
 
@@
 \end{opt_list}
-
-For the \texttt{islip}, \texttt{pim} and \texttt{select} allocators, allocation can be improved by performing multiple iterations of the algorithm; the number of iterations is controlled by the \texttt{alloc\_iters} parameter.
 

### 2008-11-20 src/allocators/loa.hpp
add Id tags
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@848 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,3 +1,4 @@
+// $Id$
 #ifndef _LOA_HPP_
 #define _LOA_HPP_
 

### 2008-11-20 src/booksim.hpp

@@ -1,3 +1,4 @@
+// $Id$
 #ifndef _BOOKSIM_HPP_
 #define _BOOKSIM_HPP_
 

### 2008-11-20 src/creditchannel.cpp

@@ -1,3 +1,4 @@
+// $Id$
 //////////////////////////////////////////////////////////////////////
 //
 //  File Name: creditchannel.cpp

### 2008-11-20 src/misc_utils.hpp

@@ -1,3 +1,4 @@
+// $Id$
 #ifndef _MISC_UTILS_HPP_
 #define _MISC_UTILS_HPP_
 

### 2008-11-20 src/networks/anynet.hpp

@@ -1,3 +1,4 @@
+// $Id$
 #ifndef _ANYNET_HPP_
 #define _ANYNET_HPP_
 

### 2008-11-20 src/networks/mecs.hpp

@@ -1,3 +1,4 @@
+// $Id$
 #ifndef _MECS_HPP_
 #define _MECS_HPP_
 

### 2008-11-20 src/networks/qtree.cpp

@@ -1,3 +1,4 @@
+// $Id$
 ////////////////////////////////////////////////////////////////////////
 //
 // QTree: A Quad-Tree Indirect Network.
@@ -8,7 +9,7 @@
 // RCS Information:
 //  $Author: jbalfour $
 //  $Date: 2007/05/17 17:14:07 $
-//  $Id: qtree.cpp,v 1.1 2007/05/17 17:14:07 jbalfour Exp $
+//  $Id$
 // 
 ////////////////////////////////////////////////////////////////////////
 

### 2008-11-20 src/networks/qtree.hpp

@@ -1,3 +1,4 @@
+// $Id$
 ////////////////////////////////////////////////////////////////////////
 //
 // QTree: A Quad-Tree Indirect Network.
@@ -8,7 +9,7 @@
 // RCS Information:
 //  $Author: jbalfour $
 //  $Date: 2007/05/17 17:14:07 $
-//  $Id: qtree.hpp,v 1.1 2007/05/17 17:14:07 jbalfour Exp $
+//  $Id$
 // 
 ////////////////////////////////////////////////////////////////////////
 

### 2008-11-20 src/networks/singlenet.cpp

@@ -1,3 +1,4 @@
+// $Id$
 /*singlenet.cpp
  *
  *single router network

### 2008-11-20 src/networks/tree4.hpp

@@ -1,3 +1,4 @@
+// $Id$
 ////////////////////////////////////////////////////////////////////////
 //
 // Tree4: Network with 64 Terminal Nodes arranged in a tree topology
@@ -8,7 +9,7 @@
 // RCS Information:
 //  $Author: jbalfour $
 //  $Date: 2007/06/26 22:49:23 $
-//  $Id: tree4.hpp,v 1.2 2007/06/26 22:49:23 jbalfour Exp $
+//  $Id$
 // 
 ////////////////////////////////////////////////////////////////////////
 

### 2008-11-20 src/power_module.hpp

@@ -1,3 +1,4 @@
+// $Id$
 #ifndef _POWER_MODULE_HPP_
 #define _POWER_MODULE_HPP_
 

### 2008-11-20 src/rng_wrapper.cpp

@@ -1,3 +1,4 @@
+// $Id$
 #include "rng.hpp"
 
 #define main rng_main

### 2008-11-20 src/routers/MECSCombiner.hpp

@@ -1,3 +1,4 @@
+// $Id$
 #ifndef _MECSCOMBINER_HPP_
 #define _MECSCOMBINER_HPP_
 

### 2008-11-20 src/routers/MECSRouter.hpp

@@ -1,3 +1,4 @@
+// $Id$
 #ifndef _MECSROUTER_HPP_
 #define _MECSROUTER_HPP_
 

### 2008-11-20 src/traffic.cpp

@@ -1,3 +1,4 @@
+// $Id$
 #include "booksim.hpp"
 #include <map>
 #include <stdlib.h>

### 2009-03-13 doc/paper.tex
First draft of paper.tex added
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1148 ce7aba0d-9f81-40a6-a495-cecae84b0515

\documentclass{article}
\usepackage{amsmath}
\usepackage{epsfig}
\usepackage[all]{xy}
\usepackage{multicol}
\usepackage{indentfirst}
\usepackage{algorithm}
\usepackage{algorithmic}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9.5in}
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\topmargin}{0in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\topskip}{0.5in}
\begin{document}
%% be sure to change the:
% Section number, Day, and time
% Date and year
% Name
% spellcheck using ``aspell -c filename'' in Unix
\begin{center}
\LARGE{\bf Parallelizing a Network Simulator}
%% PUT IN YOUR SECTION, DAY, AND TIME
%%example:
%% Section 7 - Thursday 8:30 AM\\

%% PUT IN THE DATE AND YEAR THE ASSIGNMENT IS DUE
%%example:
%% September 10, 2004
%% PUT IN YOUR NAME
\begin{center}
\Large{
Michael Bauer \hspace{1in} Nan Jiang
}
\end{center}
\Large{Department of Computer Science \\ Stanford University \\ \{mebauer, njiang37 \}@stanford.edu}
\end{center}

\begin{multicols}{2}
\begin{center}
\Large{\bf Abstract}
\end{center}
%\small{
\emph{In the context of the current multicore revolution, the need to
design an effective interconnect network has become increasingly
important.  Computer architects make use of network simulators in order to
determine the best designs and configurations.  The process by which this
is done however is often computationally intensive and requires
significant cycle time.  The fact that most simulators are single threaded
only adds to the computational overhead and renders them unable to take
advantage of new multicore processors.  This paper demonstrates how
an interconnect network simulator can be parallelized into a multithreaded
implementation.  The additional
performance gain from this parallel simulator ultimately can lead to decreased
development time and better designs as architects can sweep additional
parameter spaces and consider more configurations.}
%}
\section{Introduction}
After hitting the frequency and power walls, chip manufacturers have begun
to design and produce multicore chips that are capable of executing
multithreaded programs.  While this has the potential to yield many orders
of magnitude in performance gains, existing programs that have been
written in a single threaded programming model must be rewritten in a
parallel programming model in order to take advantage of multicore
processors.  This process of converting a program from a single threaded
programming model to a parallel programming model is a non-trivial task
that often requires restructuring the algorithm of the program as well as
how the data is organized and partioned.  In this paper we demonstate
several parallelization techniques in the process of parallelizing a
network simulator.\\
~\\
Another major ramification of the multicore revolution has been the
increased importance of interconnect network design.  As the number of
cores being placed on a chip increases, the network plays a bigger and
bigger role in minimizing the latency of communication between processors
and providing enough bandwidth to allow for the movement of large chunks
of data.  In order to make informed decisions in the process of designing
interconnect networks, computer architects make use of simulators to
observe the effect of different configurations and parameters.  While
these simulations can provide valuable insight into the best designs, they
often require sweeping a large, multi-dimensional parameter space.  In
addition to this, each simulation requires that many thousands of cycles
of operation be simulated.  This demands a large number of computational
resources in order to fully explore the design space.\\
~\\
The network simulator that we employ in this papaer was originally written
in a single threaded programming model.  This made the simulator extremely
slow and a single simulation of just a few thousand cycles would take many
minutes of compute time.  In order to speed up the process of running
simulations we have parallelized this simulator by making it
multithreaded.  In section \ref{background} we provide some background on
the simulator and its mechanics.  In section \ref{para} we describe the
means by which we parallelized the simulator.  We then describe in section
\ref{results} the results that we obtained from our parallel
implementation.  After demonstrating our performance gain and its limiting
factors we discuss how additional performance could be gained in section
\ref{disc}.  Finally, in section \ref{conc} we conclude.

\section{Simulator Background \label{background}}

\section{Parallel Implementation \label{para}}
In order to parallelize the network simulator, we first set about
identifying sources of parallelism that we could exploit.  Unlike a
mirco-architectural simulator, the network simulator doesn't have many
functional dependencies.  Each router in the network can be simulated
independently as long as the communication that occurs between the routers
is synchronized in the channels.  In this section we describe how we
partitioned the work of the network simulator and the synchronization
mechanisms that we employed in order to maintain the correctness of the simulator.

\subsection{Partitioning the Network}
The first step in parallelizing the simulator was to define a division of
work that could be spread across parallel threads with minimal amounts of
communication overhead.  Fortunately, the network simulator naturally
lends itself to this sort of partition in the sense that each router
within the network can be simulated almost entirely independent of every
other router.  In addition to this communication only occurs at the
beginning and end of every cycle of simulation when flits are either
placed into or pulled out of the shared channels.  The significant
calculation of the arbitration logic is a much larger computational that
can be performed in an entirely local manner.  While this provided us with
a simple partition of work, it left us with several important issues both
for correctness and performance.

\subsubsection{Locking Queues}
We initially assigned routers to threads by their ID numbers which are completely indpendent of
topology.  This left us no way of determining which channels would
ultimately be shared and which would be local to a thread, therefore we
were forced to place locks on all channels to ensure correctnes.  Since
the channels were intially represented as STL queues, there was no way to
allow concurrent accesses to the queues (i.e. we couldn't allow one router
to be reading and another to be writing simultaneously).  In order to
prevent this, we protected each queue with a Pthread lock.  For a thread
to access the queue, it first had to acquire the lock and then release the
lock when finished performing its operation on the queue.\\
~\\
While this approach guaranteed correctness, it also caused significant
performance overhead.  Even queues that were not shared between different
threads were having to be locked and unlocked in order to perform
operations on the queues.  In order to get around this additional
overhead, we instead began to partition the router work based upon the
topology of the network.  This therefore allowed us to know which channels
were shared and which were local to a single thread.  By doing this, we
were able to remove some of the overhead of locking when performing
operations on a channel's queue.

\subsubsection{Barriering Every Cycle}
While the difficulty of efficiently partitioning the network created
several performance issues, we also had to address the issue of the
correctness of the simulation.  By allowing each thread to operate
independently, we had created the potential for one or more threads to run
ahead of other threads in the simulation.  This means that different
threads could be performing computation on different cycles of the
simulation at the same time.  If not handled correctly, this could have
led to a correctness issue.  Consider the following example: assume that
one thread is simulating one cycle ahead of all of the other threads.  In
order for this thread to execute, it first reads in all the flits that are
in its input channels.  However, all the other threads have yet to write
to their output channels.  Therefore when the thread running ahead
performs its reads, it will observe that all its input channels are empty
and assume this is correct, when in reality, the flits simply have not
been written from the previous simulation cycle.\\
~\\
In order to successfully solve this data race, we initially place a
barrier after every cycle of the simulation.  This then mandated that
after a thread simulates a cycle, it must wait for all the other threads
in the simulator to reach the same state.  This ensures that all flits are
written into their proper channels before any thread attempts to read them
out when simulating the next cycle.  While this guarantees correctness, it
does not encourage much parallelism among the threads and adds
synchronization overhead.  In the next section we describe how we were
able to remove the barrier and the locks around the queues in order to
improve concurrency and minimize synchronization overheads.

\subsection{Lock-Free Queues}
In order to improve the concurrency of our network simulator, we decided
to incorporate the lock-free queues described in \cite{LF}.  The advantage
that these queues provide is that they allow multiple readers and writers
to the queue simulataneously.  While our simulator only requires that we
be able to have one reader and one writer, simultaneously, these queues
still allow us to remove any lock overhead from the simulation.  For
completeness we give a brief overview the implementation of lock-free
queues.

\subsubsection{Implementation}
Lock-free queues are based on the universal concurrent access algorithm
\cite{LF} that assumes that any access to the shared data structure can
see the data structure in any potential partially updated state.  The
algorithm then allows for the access to either help other accesses to
complete before actually completing itself.  While this may sound
conceptually simple, the combinatorial number of states of the data
structure and different readers and writers is astounding.\\
~\\
The lock-free queue implementation that we use begins by having a sentinel
at the head of the queue whose value is meaningless.  This eases some of
the implementation details of the queue.  Initially both the head and tail
pointers of the queue point to the sentinel.  Whenever a new item is added
to the queue, the tail pointer is moved to point to the end of the list.
Whenever an item is popped from the queue, the head is moved to point to
the new sentinel, the next item in the list.  In this case the old
sentinel can then be reclaimed.  Each element in the lock-free queue is
rapped in an AtomicReference object.  This object ensures that the any
items in the queue can only be written to atomically.  A diagram of how
this works can be seen in Figure (need figure here).\\
~\\
We initially implemented this AtomicReference object with a Pthread
lock.  While this provided a finer granularity of access to shared
objects than previously, we realized that the overhead in the number of
locks was tremendous.  Instead we relied upon an extension to GCC 4.1.2
(need citation) that provided an atomic \_\_compare\_and\_exchange operation.
This operation was significanlty less costly in terms of synchronization
overhead and was also much faster.\\
~\\
After testing our lock-free queues, we then went about replacing all of
the queues in the simulator with our new queue implementation.  While this
proved fairly straight forward, we didn't initally account for the high
overhead that the lock-free queues themselves created.  The lock-free
queues involve the creation of multiple objects and accessing the
lock-free queues involves chasing many pointers through these structures
as well as rearranging many pointers.  In some cases this overhead proved
to be extreme.  In order to remove it, we created a common interface that
was shared between the lock-free queues and the STL queue.  This then
allowed us to create dynamic instances of either a shared queue or an STL
queue depending on whether the channel was shared.  This significantly
improved our performance as we will demonstrate in the section \ref{results}.

\subsubsection{Guaranteeing Correctness Without a Barrier}
After adding the lock-free queues to our implementation, we realized that
they would provide little benefit if the barrier were not also removed to
allow threads to run ahead.  In order to remove the barrier, we had to add
an additional piece of data to each shared queue: a semaphore.  This
semaphore would be used in the following way: anytime a write operation
occurred to the queue, an up operation was performed.  Similarly, anytime
a pop operation occurred to the queue, a down operation would have to be
performed first.  We also modified the routers such that for every cycle
of simulation, they would at perform at least one up operation on every
channel that they shared with another thread even if they didn't actually
write any flits to that channel.  Note that a down operation is not able
to proceed if the semaphore equals zero, implying that no router can ever
run ahead of any of its input queues.  By adding this feature to our
lock-free queues we were able to guarantee correctness in the simulator
without a barrier operation.  This greatly increased the amount of
concurrency present in the simulation by allowing different threads to be
simulating different cycles simultaneously.

\subsection{Fusing Loops}
The last operation that we performed on our code was to fuse several loops
together.  The initial simulator implementation was written to make it
easy to understand the code and not for performance.  In some cases,
multiple independent 'for' loops were used to perform different
operations.  In several cases we moved all these operations into a single
'for' loop in order to help improve some of the temporal locality of the
code.  Not surprisingly this operation allowed the system to take
advantage of the cache hierarchy and provide additional perfromance gain.

\section{Results \label{results}}

\section{Discussion \label{disc}}

\section{Conclusion \label{conc}}

\begin{thebibliography}{99}
\bibitem{LF} Herlihy and Shavit {\em The Art of Multiprocessor Programming.} 2008.

\end{thebibliography}
\end{multicols}

\end{document}

### 2009-03-13 doc/sim.tex
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1149 ce7aba0d-9f81-40a6-a495-cecae84b0515

The interconnection network simulator, Booksim, was created by the Concurrent VLSI Architecture group (CVA) at Stanford to study network on chip and large scale computer systems. The simulator allows for a wide range of user customizations when constructing a network. The simulator supports a variety of network topologies, router micro-architecture, routing algorithms, traffic pattern, and simulation modes. The priority of the simulator is  to guarantee the correctness and accuracy of simulation, as a result the simulator is implemented in a cycle-accurate fashion. Booksim has been used to generate results for many research publications [***cite]. 
\subsection{Simulator Organization}
Booksim is written in C++ and dynamically allocates an hierarchy of objects to represent the user specified network. Figure~\ref{fig:simulator} shows the anatomy of the simulator at runtime. At the core of every network simulation is a collection of router and channel objects which forms the framework of a network. Each channel act as a FIFO with a reader and a writer router object. The router objects contain all the data structure and operations necessary to route packets through the network. Similar to networks in real life, routers in the network are mostly isolated from each other and can only communicate through network channels or simple global signals. As a result, the operations of individual routers are almost completely independent of other routers in the network. Surrounding the core network framework is the trafficmanager which acts as an interface between the user tests and the underlying network. It handles the injection and ejection of network traffic as well as gather simulations statistics for the user. \\
~\\
\begin{figure}[h]
\centering
\includegraphics[width=3in]{simulator.eps}
\caption{Simulator organization: Network is composed of routers and channels, simulations are run through the traffic manager. }
\label{fig:simulator}
\end{figure}
Figure~\ref{fig:runtime} diagrams the runtime steps of the simulator. At the highest level is the network cycle loop. Each iteration of this loop represent a cycle in the network. Within each network cycle, several steps occurs to advance the state of the network. At each step, the simulator loops through all the routers in the network to perform a specific function. As mentioned previously the operation of routers are mostly independent of each other. Therefore, there are no dependencies between each iterations of these loops. Channels are the only source of sharing between the routers 
\begin{figure}[h]
\centering
\includegraphics[width=3in]{runtime.eps}
\caption{Simulator runtime diagram. The number of iteration of the outer loop is equal to the number of simulated cycles. The number of iteration of each of the inner loops is equal the number of router in the network.}
\label{fig:runtime}
\end{figure}
\subsection{Simulator Performance}
The simulator was originally for studying numerous aspects of small to medium sized interconnection networks (less than 100 network nodes). Nearly every aspect of the network is customizable and has been modeled in detail. Performance has not been a concern. However, the group's recent project involves simulating the dragonfly network [***cite], which can easily grow to thousands or tens of thousands of nodes. Due to the organization of the simulator, the simulation runtime grows linearly with the number of network nodes and simulation time. A typical simulation of the dragonfly network involves running a networks with 10K nodes for at least 10K cycles to guarantee network steady state. Under the current simulator, this takes about 25 minutes. To generate an accurate latency-throughput plot of the network requires about 10 simulations. Furthermore, to see the effect of changes to a particular aspect of the network requires generating a latency-throughput plot each time. Such long simulation time increase the turn around time for research and limits the number of experiments that can be run reasonably.

Currently, the group utilizes program level parallelism by running multiple simulations simultaneously on the cluster. But they are still limited by the runtime of each simulation.

### 2009-03-15 doc/paper.tex
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1149 ce7aba0d-9f81-40a6-a495-cecae84b0515
Added results and conclusion sections and some files
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1150 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 \section{Simulator Background \label{background}}
+\input{sim.tex}
 
@@
 \section{Results \label{results}}
+Having completed a parallel implementation of our network simulator, our
+next step was to analyze the performance of the simulator to determine how
+much speedup it achieved.  We began first by analzying the performance of
+the different synchronization mechanisms to ensure that they held with our
+hypotheses about which permitted the most concurrency.  We then analyzed
+our results on a series of architectures to determine the scalability of
+the simulator as well as any potential bottlenecks.
+
+\subsection{Analyzing Synchronization Mechanisms}
+The first step in analyzing our parallelized network simulator was to
+determine the performance of the different synchronization mechanisms that
+we had employed at different points in development.  Figure \ref{synch}
+shows the performance of the various synchronization mechansims on a
+single node of the Hotbox cluster that consists of two Intel Xeon Quad
+cores.  We only ran experiments up to 8 threads as this is the maximum
+number of hardware thread contexts available without encountering the
+overhead of using ethernet to communicate between nodes.\\
+\begin{center}
+\includegraphics[width=2in,height=2in]{synch.eps} \\
+\small{Figure 1: This figure shows the performance of the different
+synchronization mechanisms on up to eight threads.}
+\end{center}
+From these results we can see that our hypotheses were confirmed regarding
+the performance of the various synchronization mechanisms.  Barriering
+every cycle added significant overhead to performance and ultimately
+prevented even 8 threads from extracting much parallelism at all.
+Surprisingly the same result was discovered when the barrier was removed and
+all the queues were replaced with lock-free queues.  However, once we were
+able to only implement the shared channels with lock-free queues lock-free
+queues, we could see a noticeable performance gain and even some
+additional scalability.  Finally, once we fused several of the loops being
+executed within the program, we were able to improve some of the temporal
+locality of data use and thereby achieve a little additional
+performance.\\
+~\\
+Despite our efforts however, we were only able to extract a two fold
+speedup using eight threads, far below a linear speedup.  This was very
+perplexing to us as we had consciously attempted to break up the work into
+parallel computations that required very little inter-thread
+communication.  In order to determine exactly what was causing us to lose
+performance we then proceeded to run several experiments to find the
+bottlenecks in the program.
+
+\subsection{Determining Scalability Bottlenecks}
+Having observed the somewhat poor performance of our application we then
+set about attempting to determine the reason for such poor performance.
+Our initial hypothesis was that application was memory bound based on the
+fact that there is a significant amount of data being processed in our
+simulator.  In addition to this, the fact that Pthreads is a shared memory
+programming model implies that if the working set cannot fit inside the L2
+cache, then there will be a significant amount of data being moved both to
+and from main memory.\\
+~\\
+To test our hypothesis that the simulator was memory bound, we ran the
+simulator again on the Hotbox cluster, but this time ran each thread on a
+different node.  This allowed each thread to have access to the entire
+memory bandwithd of its node, and in effect, thereby increase the total
+amount of memory bandwidth available to the entire application.  The
+results of this experiment can be seen in Figure \ref{smp}.\\
+\begin{center}
+\includegraphics[height=2in,width=2in]{smp.eps} \\
+\small{Figure 2: This figure shows the performance of the simulator on
+up to 16 threads on a simulated SMP machine.  The speedup is relative to
+the sequential simulator.}
+\end{center}
+From the result of this experiment we were able to see that bandwidth was
+not the primary issue.  This is due to the fact that the performance of
+the simulator remained poor and was even worse than the CMP version up to
+16 threads.  In addition to revealing that the poor performance was not
+due to memory bandwidth, it also revealed that communication was playing a
+large role in the performance as the increased cost of communicating via
+ethernet between nodes caused performance to be even worse than the CMP
+configuration.\\
+~\\
+After realizing that our program was not memory bound, we wanted to see if
+the size of the problem had any effect on the speedup.  We tested this by
+plotting the speedup of a four threaded run on a given network size over the
+sequential time for simulating the same size network on the Hotbox
+cluster.  The results can be seen in Figure \ref{probsize}.
+\begin{center}
+\includegraphics[height=2in,width=2in]{probsize.eps}
+\small{Figure 3: This figure shows the speedup versus network size of a
+four threaded run of the parallel simulator on the Hotbox cluster.}
+\end{center}
+From this figure we could see that essentially the speedup was independent
+of problem size.  This meant that regardless of the size of the network
+simulated, the four-threaded version ultimately was constrained by the
+same relative to the sequential version.  The fact that the speedup
+attained was significantly less than a linear speedup of 4 indicated that
+the threads were not compute bound.  Instead we hypothesized that false
+sharing was causing the four threads to thrash each others' caches.\\
+~\\
+To confirm our diagnosis of false sharing, we decided to run the simulator
+on the both the Niagara T2 and T2+ machines.  Since these machines were
+better designed to deal with long latency loads and stores, our hope was
+that they would be able to mitigate the effects of false sharing to some
+extend.  The results from this experiment can be seen in figure
+\ref{niagara}.
+\begin{center}
+\includegraphics[height=2in,width=2in]{niagara.eps}
+\small{Figure 4: This figure shows the speedup of the parallel simulator
+up to 64 threads on the Niagara T2 and T2+ machines.}
+\end{center}
+We noticed that in this experiment, both the T2 and T2+ were able to
+achieve some speedup to a certain number of threads, but then performance
+began to fall off.  We determined that once the threads' working set sizes
+began to interfere with each other in the shared L2 cache is when the
+performance begins to degrade and couldn't offer us much insight into the
+false any potential false sharing that was occurring when run on other
+hardware platforms.
 
@@
 \section{Conclusion \label{conc}}
+We have developed a parallelized version of a interconnect network
+simulator that is capable of achieving some speedup on both CMP and SMP
+machines.  We outlined the basics of the simulator and the various
+techniques that we used in order to extract concurrency from the
+simulator.  We then demonstrated the performance of our parallel simulator
+on several different hardware platforms.  Investigtions showed that the
+simulator is prone to sufferring from false sharing and extremem
+communication overheads.  Despite these bottlenecks we were able to
+achieve up to a 3X performance which can significantly lower the
+simulation time required for running future experiments.
 
 \begin{thebibliography}{99}
-\bibitem{LF} Herlihy and Shavit {\em The Art of Multiprocessor Programming.} 2008.
+\bibitem{LF} Herlihy, Maurice and Shavit, Nir {\em The Art of Multiprocessor Programming.} 2008.
 

### 2009-03-15 doc/paper.tex
changes
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1151 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 design an effective interconnect network has become increasingly
-important.  Computer architects make use of network simulators in order to
-determine the best designs and configurations.  The process by which this
+vital.  Computer architects make use of network simulators in order to
+determine the best network designs and configurations.  The process by which this
 is done however is often computationally intensive and requires
-significant cycle time.  The fact that most simulators are single threaded
-only adds to the computational overhead and renders them unable to take
+significant run time.  The fact that most simulators are single threaded
+only adds to the computational bottleneck and renders them unable to take
 advantage of new multicore processors.  This paper demonstrates how
@@
 to design and produce multicore chips that are capable of executing
-multithreaded programs.  While this has the potential to yield many orders
-of magnitude in performance gains, existing programs that have been
+multithreaded programs.  While this has greatly increased the computation power of the processors, existing programs that have been
 written in a single threaded programming model must be rewritten in a
-parallel programming model in order to take advantage of multicore
-processors.  This process of converting a program from a single threaded
+parallel programming model in order to take advantage of the additional cores.  This process of converting a program from a single threaded
 programming model to a parallel programming model is a non-trivial task
 that often requires restructuring the algorithm of the program as well as
-how the data is organized and partioned.  In this paper we demonstate
+how the data is organized and partitioned.  In this paper we demonstrate
 several parallelization techniques in the process of parallelizing a
@@
 increased importance of interconnect network design.  As the number of
-cores being placed on a chip increases, the network plays a bigger and
+cores being placed on a chip increases, the processor interconnect plays a bigger and
 bigger role in minimizing the latency of communication between processors
 and providing enough bandwidth to allow for the movement of large chunks
-of data.  In order to make informed decisions in the process of designing
+of data.
+Traditional bus architecture is no longer the appropriate medium of communication for high number of core.
+Even more recent single crossbar communication hubs will also not be sufficient in the future, since the complexity of a single crossbar grows quadratically with the number of cores.
+Interconnection network can provide the scalable and efficient communication
+infrastructures for future multicore processors. \\
+~\\
+In order to make informed decisions in the process of designing
 interconnect networks, computer architects make use of simulators to
-observe the effect of different configurations and parameters.  While
+observe the effect of different network configurations and parameters.  While
 these simulations can provide valuable insight into the best designs, they
 often require sweeping a large, multi-dimensional parameter space.  In
-addition to this, each simulation requires that many thousands of cycles
-of operation be simulated.  This demands a large number of computational
+addition, each simulation requires that many thousands of cycles
+of network operation to be simulated.  This demands a large number of computational
 resources in order to fully explore the design space.\\
@@
 In order to parallelize the network simulator, we first set about
-identifying sources of parallelism that we could exploit.  Unlike a
-mirco-architectural simulator, the network simulator doesn't have many
-functional dependencies.  Each router in the network can be simulated
+identifying sources of parallelism that we could exploit.  As mentioned previous, the routers in a network operate relatively autonomously.   Each router in the network can be simulated
 independently as long as the communication that occurs between the routers
@@
 beginning and end of every cycle of simulation when flits are either
-placed into or pulled out of the shared channels.  The significant
+written or read from the shared channels.  The significant
 calculation of the arbitration logic is a much larger computational that
@@
 
-\subsubsection{Locking Queues}
-We initially assigned routers to threads by their ID numbers which are completely indpendent of
-topology.  This left us no way of determining which channels would
+\subsubsection{Locking Channels}
+We initially assigned routers to threads by their ID numbers which are completely independent of
+topology.  This left us no easy way of determining which channels would
 ultimately be shared and which would be local to a thread, therefore we
-were forced to place locks on all channels to ensure correctnes.  Since
-the channels were intially represented as STL queues, there was no way to
+were forced to place locks on all channels to ensure correctness.  Since
+the channels were initially represented as STL queues, there was no way to
 allow concurrent accesses to the queues (i.e. we couldn't allow one router
@@
 operations on a channel's queue.
+However, this method adds an additional layer of complexity for the user.
+For each topology in the simulator, we need to write a network partitioning algorithm for that topology.
+Ideally we would like a generalized algorithm that would discover the optimal partitioning to minimize locking given any network.
+But this is not part of the scope of this project and has not been done.
+(***LOL NP COMPLETE??)
 
@@
 led to a correctness issue.  Consider the following example: assume that
-one thread is simulating one cycle ahead of all of the other threads.  In
+one thread is simulating one cycle ahead of all of the other threads and the network channel latency is also one cycle.  In
 order for this thread to execute, it first reads in all the flits that are
 in its input channels.  However, all the other threads have yet to write
-to their output channels.  Therefore when the thread running ahead
+to their output channels from the previous cycle.  Therefore when the thread running ahead
 performs its reads, it will observe that all its input channels are empty
@@
 after a thread simulates a cycle, it must wait for all the other threads
-in the simulator to reach the same state.  This ensures that all flits are
+in the simulator to reach the same point.  This ensures that all flits are
 written into their proper channels before any thread attempts to read them
 out when simulating the next cycle.  While this guarantees correctness, it
-does not encourage much parallelism among the threads and adds
-synchronization overhead.  In the next section we describe how we were
+can add significant
+synchronization overhead.
+Transient load imbalances in the network may occur even under uniform random traffic pattern, this will the cause some threads to run slower temporarily.
+The barrier will amplify these transient load imbalance since the simulator will run as fast as the slowest thread.
+In the next section we describe how we were
 able to remove the barrier and the locks around the queues in order to
@@
 that these queues provide is that they allow multiple readers and writers
-to the queue simulataneously.  While our simulator only requires that we
+to the queue simultaneously.  While our simulator only requires that we
 be able to have one reader and one writer, simultaneously, these queues
@@
 sentinel can then be reclaimed.  Each element in the lock-free queue is
-rapped in an AtomicReference object.  This object ensures that the any
+wrapped in an AtomicReference object.  This object ensures that the any
 items in the queue can only be written to atomically.  A diagram of how
@@
 (need citation) that provided an atomic \_\_compare\_and\_exchange operation.
-This operation was significanlty less costly in terms of synchronization
+This operation was significantly less costly in terms of synchronization
 overhead and was also much faster.\\
 ~\\
-After testing our lock-free queues, we then went about replacing all of
+After testing our lock-free queues, we replaced all of
 the queues in the simulator with our new queue implementation.  While this
-proved fairly straight forward, we didn't initally account for the high
+proved fairly straight forward, we didn't initially account for the high
 overhead that the lock-free queues themselves created.  The lock-free
@@
 \subsection{Fusing Loops}
-The last operation that we performed on our code was to fuse several loops
-together.  The initial simulator implementation was written to make it
-easy to understand the code and not for performance.  In some cases,
-multiple independent 'for' loops were used to perform different
-operations.  In several cases we moved all these operations into a single
+The last operation that we performed on our code was to the inner loops in figure~\ref{fig:runtime} together.  The initial simulator implementation was written to make it
+easy to understand the code and not for performance.  We moved all these operations into a single
 'for' loop in order to help improve some of the temporal locality of the
 code.  Not surprisingly this operation allowed the system to take
-advantage of the cache hierarchy and provide additional perfromance gain.
+advantage of the cache hierarchy and provide additional performance gain.
 
@@
 next step was to analyze the performance of the simulator to determine how
-much speedup it achieved.  We began first by analzying the performance of
+much speedup it achieved.  We began first by analyzing the performance of
 the different synchronization mechanisms to ensure that they held with our
-hypotheses about which permitted the most concurrency.  We then analyzed
+hypotheses about which method permitted the most concurrency.  We then analyzed
 our results on a series of architectures to determine the scalability of
@@
 all the queues were replaced with lock-free queues.  However, once we were
-able to only implement the shared channels with lock-free queues lock-free
-queues, we could see a noticeable performance gain and even some
+able to only implement the shared channels with lock-free queues , we could see a noticeable performance gain and even some
 additional scalability.  Finally, once we fused several of the loops being
@@
 different node.  This allowed each thread to have access to the entire
-memory bandwithd of its node, and in effect, thereby increase the total
+memory bandwidth of its node, and in effect, thereby increase the total
 amount of memory bandwidth available to the entire application.  The
@@
 began to interfere with each other in the shared L2 cache is when the
-performance begins to degrade and couldn't offer us much insight into the
-false any potential false sharing that was occurring when run on other
+performance begins to degrade and couldn't offer us much insight into  any potential false sharing that was occurring when run on other
 hardware platforms.
@@
 \section{Discussion \label{disc}}
+\input{disc.tex}
 
@@
 on several different hardware platforms.  Investigtions showed that the
-simulator is prone to sufferring from false sharing and extremem
+simulator is prone to sufferring from false sharing and extreme
 communication overheads.  Despite these bottlenecks we were able to
@@
 
-\begin{thebibliography}{99}
-\bibitem{LF} Herlihy, Maurice and Shavit, Nir {\em The Art of Multiprocessor Programming.} 2008.
+%\begin{thebibliography}{99}
+%\bibitem{LF} Herlihy, Maurice and Shavit, Nir {\em The Art of Multiprocessor Programming.} 2008.
 
-\end{thebibliography}
+%\end{thebibliography}
+
+\bibliographystyle{IEEEtran}
+\bibliography{paper}
 \end{multicols}

### 2009-03-15 doc/sim.tex
changes
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1151 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 The priority of the simulator is  to guarantee the correctness and accuracy of simulation, as a result the simulator is implemented in a cycle-accurate fashion.
-Booksim has been used to generate results for many research publications [***cite].
+Booksim has been used to generate results for many research publications \cite{cdragon,KimISCA07,PPIN}.
 \subsection{Simulator Organization}
@@
 The router objects contain all the data structure and operations necessary to route packets through the network.
-Similar to networks in real life, routers in the network are mostly isolated from each other and can only communicate through network channels or simple global signals.
+Similar to networks in real life, routers are mostly isolated from each other and can only communicate through channels or simple global signals.
 As a result, the operations of individual routers are almost completely independent of other routers in the network.
-Surrounding the core network framework is the trafficmanager which acts as an interface between the user tests and the underlying network.
-It handles the injection and ejection of network traffic as well as gather simulations statistics for the user.
+Surrounding the core network framework is the traffic manager which acts as an interface between the user tests and the underlying network.
+It handles the injection and ejection of traffic as well as gather simulations statistics for the user.
 \\
 ~\\
-\begin{figure}[h]
+\begin{figure*}[h]
 \centering
@@
 \label{fig:simulator}
-\end{figure}
+\end{figure*}
 Figure~\ref{fig:runtime} diagrams the runtime steps of the simulator.
@@
 Channels are the only source of sharing between the routers
-\begin{figure}[h]
+\begin{figure*}[h]
 \centering
@@
 \label{fig:runtime}
-\end{figure}
+\end{figure*}
 \subsection{Simulator Performance}
-The simulator was originally for studying numerous aspects of small to medium sized interconnection networks (less than 100 network nodes).
+The simulator was originally designed for studying numerous aspects of small to medium sized interconnection networks (less than 100 network nodes).
 Nearly every aspect of the network is customizable and has been modeled in detail.
 Performance has not been a concern.
-However, the group's recent project involves simulating the dragonfly network [***cite], which can easily grow to thousands or tens of thousands of nodes.
+However, the group's recent project involves simulating the dragonfly network \cite{cdragon}, which can easily grow to thousands or tens of thousands of nodes.
 Due to the organization of the simulator, the simulation runtime grows linearly with the number of network nodes and simulation time.
-A typical simulation of the dragonfly network involves running a networks with 10K nodes for at least 10K cycles to guarantee network steady state.
+A typical simulation of the dragonfly network involves running a networks with 10K nodes for at least 10K cycles to capture network steady state.
 Under the current simulator, this takes about 25 minutes.
@@
 Such long simulation time increase the turn around time for research and limits the number of experiments that can be run reasonably.
+Currently, the group utilizes program level parallelism by running multiple simulations simultaneously on the cluster.
+But they are still limited by the runtime of each simulation.
 
-Currently, the group utilizes program level parallelism by running multiple simulations simultaneously on the cluster. But they are still limited by the runtime of each simulation.
+Previously we have profiled the simulator to determine the program characteristics and hotspots.
+The program is composed of mostly integer operations.
+The only sources of floating calculation are for random packet generation and statistics collection.
+The program is control flow intensive, there are numerous branches inside of long running loops to determine the flow of packets and state of the routers.
+The program is also relatively memory demanding, since each network component is created as an object and network packets objects are dynamically allocated and deallocated as they enter and exit the network.
+In the typical dragonfly network modeled with realistic channel latencies, there is potentially hundreds of thousands of packets in flight simultaneously.
+Runtime profiling shows that 95\% of program time is spent in the routing step in figure~\ref{fig:runtime}.

### 2009-03-16 doc/paper.tex
sitll writing the discussion section
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1152 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 \begin{center}
-\includegraphics[width=2in,height=2in]{synch.eps} \\
+\includegraphics[width=3in,height=3in]{synch.eps} \\
 \small{Figure 1: This figure shows the performance of the different
@@
 \begin{center}
-\includegraphics[height=2in,width=2in]{smp.eps} \\
+\includegraphics[height=3in,width=3in]{smp.eps} \\
 \small{Figure 2: This figure shows the performance of the simulator on
@@
 \begin{center}
-\includegraphics[height=2in,width=2in]{probsize.eps}
+\includegraphics[height=3in,width=3in]{probsize.eps}
 \small{Figure 3: This figure shows the speedup versus network size of a
@@
 \begin{center}
-\includegraphics[height=2in,width=2in]{niagara.eps}
+\includegraphics[height=3in,width=3in]{niagara.eps}
 \small{Figure 4: This figure shows the speedup of the parallel simulator
@@
 simulator.  We then demonstrated the performance of our parallel simulator
-on several different hardware platforms.  Investigtions showed that the
-simulator is prone to sufferring from false sharing and extreme
+on several different hardware platforms.  Investigations showed that the
+simulator is prone to suffering from false sharing and extreme
 communication overheads.  Despite these bottlenecks we were able to

### 2009-03-16 doc/disc.tex
discussion
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1153 ce7aba0d-9f81-40a6-a495-cecae84b0515

\subsection{Static Network Partition}
The current implementation divides the network into equal partitions among the threads and seeks to minimize computation. This organization works well when the traffic pattern is balanced across the partitions. But hotspot traffic patterns which concentrate the packet into a particular section of the network will causes imbalance in the amount of work handled by each thread. Even with the shared channel semaphore and the removal of cycle barrier, the lightly loaded threads will almost always be stalled waiting for the hotspot threads. Fortunately, common network traffic patterns are balanced and network routing algorithm are usually designed to reduce hotspots and bottlenecks in the network. Nevertheless, it is not a guarantee that all an equally partitioned network will give equal work to all threads. Therefore, future improvements for the simulator should include dynamically partitioning of the network among the threads. This can be implemented by changing the current loop based iteration of network components into a work queue based organization. A work queue can be created for each network partition and each queue entry record the network component to be operated on and the network cycle associated with the operation. After operating on a component, the component is reinserted into the queue and the associated cycle is incremented. Under this organization, threads that have stalled ahead can pull work from other thread work queue and help to finish the work in other network partitions.

\subsection{Programming Model}
Given the underwhelming speedup that we obtained through parallelization and the possibility of excessive false sharing in the program, this has lead us to think that maybe PThread is not the appropriate programming model for this simulator. As have stated through out the report, the only points of sharing between threads are the channels that bridge routers running under different threads. No other source of true sharing should exist in the simulator. But with the shared memory model of PThreads, false sharing due to the organization of data structure can easily occur. Since we know the exact points of data sharing and since the smart partitioning of the networks should keep the amount of inter-thread communication small, perhaps message passing programming models such as MPI will offer more scalability for the parallel simulator. Shared channel objects can be written on top of MPI's send and receive interfaces, and many packets can be bundled together to reduce messaging overhead.

### 2009-03-16 doc/paper.tex
title lol
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1154 ce7aba0d-9f81-40a6-a495-cecae84b0515
Fixed figure dependencies after removing multicols.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1155 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
-\documentclass{article}
+\documentclass[twocolumn]{article}
 \usepackage{amsmath}
@@
 \usepackage[all]{xy}
-\usepackage{multicol}
+%\usepackage{multicol}
 \usepackage{indentfirst}
@@
 \setlength{\topskip}{0.5in}
-\begin{document}
 %% be sure to change the:
@@
 % spellcheck using ``aspell -c filename'' in Unix
-\begin{center}
-\LARGE{\bf Parallelizing a Network Simulator}
+%\begin{center}
+%\LARGE{\bf Parallelizing a Network Simulator}
 %% PUT IN YOUR SECTION, DAY, AND TIME
@@
 %% PUT IN YOUR NAME
-\begin{center}
-\Large{
+%\begin{center}
+%\Large{
+%Michael Bauer \hspace{1in} Nan Jiang
+%}
+%\end{center}
+%\Large{Department of Computer Science \\ Stanford University \\ \{mebauer, njiang37 %\}@stanford.edu}
+%\end{center}
+
+\title{\LARGE{\bf Parallelizing a Network Simulator}}
+\author{
 Michael Bauer \hspace{1in} Nan Jiang
 }
-\end{center}
-\Large{Department of Computer Science \\ Stanford University \\ \{mebauer, njiang37 \}@stanford.edu}
-\end{center}
 
-\begin{multicols}{2}
-\begin{center}
-\Large{\bf Abstract}
-\end{center}
+\date{\Large{Department of Computer Science \\ Stanford University \\ \{mebauer, njiang37 \}@stanford.edu}}
+%\maketitle
+
+%\begin{multicols}{2}
+%\begin{center}
+%\Large{\bf Abstract}
+%\end{center}
 %\small{
-\emph{In the context of the current multicore revolution, the need to
+%\emph{
+\begin{document}
+\maketitle
+\abstract{ 
+In the context of the current multicore revolution, the need to
 design an effective interconnect network has become increasingly
@@
 overhead of using ethernet to communicate between nodes.\\
-\begin{center}
-\includegraphics[width=3in,height=3in]{synch.eps} \\
-\small{Figure 1: This figure shows the performance of the different
+\begin{figure}[h]
+\centering
+\includegraphics[width=3in,height=3in]{synch.eps}
+\caption{\label{org} This figure shows the performance of the different
 synchronization mechanisms on up to eight threads.}
-\end{center}
+\end{figure}
 From these results we can see that our hypotheses were confirmed regarding
@@
 results of this experiment can be seen in Figure \ref{smp}.\\
-\begin{center}
-\includegraphics[height=3in,width=3in]{smp.eps} \\
-\small{Figure 2: This figure shows the performance of the simulator on
+\begin{figure}[h]
+\centering
+\includegraphics[height=3in,width=3in]{smp.eps}
+\caption{\label{smp}This figure shows the performance of the simulator on
 up to 16 threads on a simulated SMP machine.  The speedup is relative to
 the sequential simulator.}
-\end{center}
+\end{figure}
 From the result of this experiment we were able to see that bandwidth was
@@
 cluster.  The results can be seen in Figure \ref{probsize}.
-\begin{center}
+\begin{figure}[h]
+\centering
 \includegraphics[height=3in,width=3in]{probsize.eps}
-\small{Figure 3: This figure shows the speedup versus network size of a
+\caption{\label{probsize}This figure shows the speedup versus network size of a
 four threaded run of the parallel simulator on the Hotbox cluster.}
-\end{center}
+\end{figure}
 From this figure we could see that essentially the speedup was independent
@@
 \ref{niagara}.
-\begin{center}
+\begin{figure}[h]
+\centering
 \includegraphics[height=3in,width=3in]{niagara.eps}
-\small{Figure 4: This figure shows the speedup of the parallel simulator
+\caption{\label{niagara}This figure shows the speedup of the parallel simulator
 up to 64 threads on the Niagara T2 and T2+ machines.}
-\end{center}
+\end{figure}
 We noticed that in this experiment, both the T2 and T2+ were able to
@@
 \bibliography{paper}
-\end{multicols}
+%\end{multicols}
 

### 2009-03-16 doc/disc.tex
First revision of paper done.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1156 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 This organization works well when the traffic pattern is balanced across the partitions.
-But hotspot traffic patterns which concentrate the packet into a particular section of the network will causes imbalance in the amount of work handled by each thread.
-Even with the shared channel semaphore and the removal of cycle barrier, the lightly loaded threads will almost always be stalled waiting for the hotspot threads.
+However, hotspot traffic patterns which concentrate the packets into a particular section of the network will cause imbalances in the amount of work handled by each thread.
+Even with the shared channel semaphore and the removal of the cycle barrier, the lightly loaded threads will almost always be stalled waiting for the hotspot threads.
 Fortunately, common network traffic patterns are balanced and network routing algorithm are usually designed to reduce hotspots and bottlenecks in the network.
-Nevertheless, it is not a guarantee that all an equally partitioned network will give equal work to all threads.
-Therefore, future improvements for the simulator should include dynamically partitioning of the network among the threads.
+Nevertheless, it is not a guarantee that an equally partitioned network will give equal work to all threads.
+Therefore, future improvements for the simulator should include dynamic partitioning of the network among the threads.
 This can be implemented by changing the current loop based iteration of network components into a work queue based organization.
@@
 After operating on a component, the component is reinserted into the queue and the associated cycle is incremented.
-Under this organization, threads that have stalled ahead can pull work from other thread work queue and help to finish the work in other network partitions.
+Under this organization, threads that have stalled ahead can pull work from other threads' work queues and help to finish the work in other network partitions.
 
 \subsection{Programming Model}
-Given the underwhelming speedup that we obtained through parallelization and the possibility of excessive false sharing in the program, this has lead us to think that maybe PThread is not the appropriate programming model for this simulator.
-As have stated through out the report, the only points of sharing between threads are the channels that bridge routers running under different threads.
+Given the underwhelming speedup that we obtained through parallelization and the possibility of excessive false sharing in the program, this has lead us to think that maybe PThreads is not the most appropriate programming model for this simulator.
+As we have stated through out the report, the only points of sharing between threads are the channels that bridge routers running under different threads.
 No other source of true sharing should exist in the simulator.

### 2009-03-16 doc/paper.tex
First revision of paper done.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1156 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 development time and better designs as architects can sweep additional
-parameter spaces and consider more configurations.}
+parameter spaces and consider more exotic configurations.}
 %}
@@
 ~\\
-Another major ramification of the multicore revolution has been the
+In addition to the need for parallel programs, another major ramification of the multicore revolution has been the
 increased importance of interconnect network design.  As the number of
@@
 of data.
-Traditional bus architecture is no longer the appropriate medium of communication for high number of core.
-Even more recent single crossbar communication hubs will also not be sufficient in the future, since the complexity of a single crossbar grows quadratically with the number of cores.
+The traditional bus architecture is no longer the appropriate medium of communication for large number of cores.
+Even more recently, single crossbar communication hubs will also not be sufficient in the future since the complexity of a single crossbar grows quadratically with the number of cores.
 Interconnection network can provide the scalable and efficient communication
@@
 addition, each simulation requires that many thousands of cycles
-of network operation to be simulated.  This demands a large number of computational
+of network operation be simulated.  This demands a large number of computational
 resources in order to fully explore the design space.\\
 ~\\
-The network simulator that we employ in this papaer was originally written
+The network simulator that we employ in this paper was originally written
 in a single threaded programming model.  This made the simulator extremely
@@
 within the network can be simulated almost entirely independent of every
-other router.  In addition to this communication only occurs at the
+other router.  In addition to this, communication only occurs at the
 beginning and end of every cycle of simulation when flits are either
 written or read from the shared channels.  The significant
-calculation of the arbitration logic is a much larger computational that
+calculation of the arbitration logic is a much larger computation that
 can be performed in an entirely local manner.  While this provided us with
@@
 to be reading and another to be writing simultaneously).  In order to
-prevent this, we protected each queue with a Pthread lock.  For a thread
+prevent this, we protected each queue with a PThread lock.  For a thread
 to access the queue, it first had to acquire the lock and then release the
@@
 overhead, we instead began to partition the router work based upon the
-topology of the network.  This therefore allowed us to know which channels
+topology of the network.  This allowed us to know which channels
 were shared and which were local to a single thread.  By doing this, we
@@
 For each topology in the simulator, we need to write a network partitioning algorithm for that topology.
-Ideally we would like a generalized algorithm that would discover the optimal partitioning to minimize locking given any network.
-But this is not part of the scope of this project and has not been done.
-(***LOL NP COMPLETE??)
+Ideally we would like a generalized algorithm that would discover the optimal partitioning to minimize locking given any network, but this is not within the scope of this project and is a very difficult (NP-Complete) problem by itself.
 
@@
 ~\\
-In order to successfully solve this data race, we initially place a
+In order to successfully solve this data race, we initially placed a
 barrier after every cycle of the simulation.  This then mandated that
@@
 synchronization overhead.
-Transient load imbalances in the network may occur even under uniform random traffic pattern, this will the cause some threads to run slower temporarily.
-The barrier will amplify these transient load imbalance since the simulator will run as fast as the slowest thread.
+Transient load imbalances in the network may occur even under uniform random traffic patterns and this will the cause some threads to run slower temporarily.
+The barrier will amplify these transient load imbalances since the simulator will only run as fast as the slowest thread.
 In the next section we describe how we were
@@
 that these queues provide is that they allow multiple readers and writers
-to the queue simultaneously.  While our simulator only requires that we
-be able to have one reader and one writer, simultaneously, these queues
+to access the queue simultaneously.  While our simulator only requires that we
+be able to have one reader and one writer simultaneously, these queues
 still allow us to remove any lock overhead from the simulation.  For
@@
 \subsubsection{Implementation}
-Lock-free queues are based on the universal concurrent access algorithm
+Lock-free queues are based on the universal lock-free algorithm
 \cite{LF} that assumes that any access to the shared data structure can
 see the data structure in any potential partially updated state.  The
-algorithm then allows for the access to either help other accesses to
+algorithm then allows for the access to help other accesses to
 complete before actually completing itself.  While this may sound
@@
 ~\\
-We initially implemented this AtomicReference object with a Pthread
+We initially implemented this AtomicReference object with a PThread
 lock.  While this provided a finer granularity of access to shared
@@
 locks was tremendous.  Instead we relied upon an extension to GCC 4.1.2
-(need citation) that provided an atomic \_\_compare\_and\_exchange operation.
+that provided an atomic \_\_compare\_and\_exchange operation.
 This operation was significantly less costly in terms of synchronization
@@
 After testing our lock-free queues, we replaced all of
-the queues in the simulator with our new queue implementation.  While this
+the queues in the simulator with our lock-free queue implementation.  While this
 proved fairly straight forward, we didn't initially account for the high
@@
 allowed us to create dynamic instances of either a shared queue or an STL
-queue depending on whether the channel was shared.  This significantly
-improved our performance as we will demonstrate in the section \ref{results}.
+queue depending on whether the channel was shared or not.  This significantly
+improved our performance as we will demonstrate in section \ref{results}.
 
@@
 they would provide little benefit if the barrier were not also removed to
-allow threads to run ahead.  In order to remove the barrier, we had to add
+allow threads to run ahead.  In order to remove the barrier and still ensure correctness of the simulator, we had to add
 an additional piece of data to each shared queue: a semaphore.  This
@@
 performed first.  We also modified the routers such that for every cycle
-of simulation, they would at perform at least one up operation on every
+of simulation, they would perform at least one up operation on every
 channel that they shared with another thread even if they didn't actually
@@
 \subsection{Fusing Loops}
-The last operation that we performed on our code was to the inner loops in figure~\ref{fig:runtime} together.  The initial simulator implementation was written to make it
+The last operation that we performed on our code was to fuse the inner loops in figure~\ref{fig:runtime} together.  The initial simulator implementation was written to make it
 easy to understand the code and not for performance.  We moved all these operations into a single
@@
 code.  Not surprisingly this operation allowed the system to take
-advantage of the cache hierarchy and provide additional performance gain.
+advantage of the cache hierarchy and provided additional performance gain.
 
@@
 we had employed at different points in development.  Figure \ref{synch}
-shows the performance of the various synchronization mechansims on a
+shows the performance of the various synchronization mechanisms on a
 single node of the Hotbox cluster that consists of two Intel Xeon Quad
@@
 \includegraphics[width=3in,height=3in]{synch.eps}
-\caption{\label{org} This figure shows the performance of the different
-synchronization mechanisms on up to eight threads.}
+\caption{\label{synch} This figure shows the performance of the different
+synchronization mechanisms on up to eight threads on the Hotbox cluster.}
 \end{figure}
+~\\
 From these results we can see that our hypotheses were confirmed regarding
@@
 
-\subsection{Determining Scalability Bottlenecks}
+\subsection{Locating Scalability Bottlenecks}
 Having observed the somewhat poor performance of our application we then
@@
 fact that there is a significant amount of data being processed in our
-simulator.  In addition to this, the fact that Pthreads is a shared memory
+simulator.  In addition to this, the fact that PThreads is a shared memory
 programming model implies that if the working set cannot fit inside the L2
@@
 \caption{\label{smp}This figure shows the performance of the simulator on
-up to 16 threads on a simulated SMP machine.  The speedup is relative to
+up to 16 threads on a simulated SMP machine on the Hotbox cluster.  The speedup is relative to
 the sequential simulator.}
@@
 simulated, the four-threaded version ultimately was constrained by the
-same relative to the sequential version.  The fact that the speedup
+same effect as the sequential version.  The fact that the speedup
 attained was significantly less than a linear speedup of 4 indicated that
@@
 that they would be able to mitigate the effects of false sharing to some
-extend.  The results from this experiment can be seen in figure
+extent.  The results from this experiment can be seen in figure
 \ref{niagara}.

### 2009-03-16 doc/sim.tex
Fixed figure dependencies after removing multicols.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1155 ce7aba0d-9f81-40a6-a495-cecae84b0515
First revision of paper done.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1156 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 The simulator allows for a wide range of user customizations when constructing a network.
-The simulator supports a variety of network topologies, router micro-architecture, routing algorithms, traffic pattern, and simulation modes.
-The priority of the simulator is  to guarantee the correctness and accuracy of simulation, as a result the simulator is implemented in a cycle-accurate fashion.
+The simulator supports a variety of network topologies, router micro-architectures, routing algorithms, traffic pattern, and simulation modes.
+The priority of the simulator is  to guarantee the correctness and accuracy of simulation, and as a result the simulator is implemented in a cycle-accurate fashion.
 Booksim has been used to generate results for many research publications \cite{cdragon,KimISCA07,PPIN}.
 \subsection{Simulator Organization}
-Booksim is written in C++ and dynamically allocates an hierarchy of objects to represent the user specified network.
+Booksim is written in C++ and dynamically allocates a hierarchy of objects to represent the user specified network.
 Figure~\ref{fig:simulator} shows the anatomy of the simulator at runtime.
 At the core of every network simulation is a collection of router and channel objects which forms the framework of a network.
-Each channel act as a FIFO with a reader and a writer router object.
+Each channel acts as a FIFO with a reader and a writer router object.
 The router objects contain all the data structure and operations necessary to route packets through the network.
@@
 Surrounding the core network framework is the traffic manager which acts as an interface between the user tests and the underlying network.
-It handles the injection and ejection of traffic as well as gather simulations statistics for the user.
+It handles the injection and ejection of traffic as well as gathering simulations statistics for the user.
 \\
 ~\\
-\begin{figure*}[h]
+\begin{figure}[h]
 \centering
 \includegraphics[width=3in]{simulator.eps}
-\caption{Simulator organization: Network is composed of routers and channels, simulations are run through the traffic manager. }
+\caption{Simulator organization: Network is composed of routers and channels and simulations are run through the traffic manager. }
 \label{fig:simulator}
-\end{figure*}
+\end{figure}
 Figure~\ref{fig:runtime} diagrams the runtime steps of the simulator.
 At the highest level is the network cycle loop.
-Each iteration of this loop represent a cycle in the network.
+Each iteration of this loop represents a cycle in the network.
 Within each network cycle, several steps occurs to advance the state of the network.
@@
 Channels are the only source of sharing between the routers
-\begin{figure*}[h]
+\begin{figure}[h]
 \centering
@@
 \caption{Simulator runtime diagram.
-The number of iteration of the outer loop is equal to the number of simulated cycles.
-The number of iteration of each of the inner loops is equal the number of router in the network.}
+The number of iterations of the outer loop is equal to the number of simulated cycles.
+The number of iterations of each of the inner loops is equal the number of routers in the network.}
 \label{fig:runtime}
-\end{figure*}
+\end{figure}
 \subsection{Simulator Performance}
@@
 Due to the organization of the simulator, the simulation runtime grows linearly with the number of network nodes and simulation time.
-A typical simulation of the dragonfly network involves running a networks with 10K nodes for at least 10K cycles to capture network steady state.
+A typical simulation of the dragonfly network involves running a networks with 10K nodes for at least 10K cycles to capture the network's steady state.
 Under the current simulator, this takes about 25 minutes.
@@
 The only sources of floating calculation are for random packet generation and statistics collection.
-The program is control flow intensive, there are numerous branches inside of long running loops to determine the flow of packets and state of the routers.
+The program is control flow intensive: there are numerous branches inside of long running loops to determine the flow of packets and state of the routers.
 The program is also relatively memory demanding, since each network component is created as an object and network packets objects are dynamically allocated and deallocated as they enter and exit the network.

### 2009-03-16 doc/paper.tex
Added a figure for the lfqueue implementation.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1157 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 items in the queue can only be written to atomically.  A diagram of how
-this works can be seen in Figure (need figure here).\\
-~\\
+this works can be seen in Figure \ref{lfqueue}.
+\begin{figure}[h]
+\centering
+\includegraphics[width=3in, height=2.25in]{lfqueue.eps}
+\caption{This figure demonstrates the layout of the objects associated with a lock-free queue of length 2 (the sentinel value does not count to towards the length of the list).
+Note that the pointer to the next element in the list, does not point directly to next list node, but instead to the wrapper AtomicReference class.
+Also note that the sentinel is not an actual maintained pointer, but will always point to the same AtomicReference as the Head Node.
+It is included in this diagram for clarity.}
+\label{lfqueue}
+\end{figure}
 We initially implemented this AtomicReference object with a PThread

### 2009-03-16 doc/disc.tex
I am good with it
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1158 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 However, hotspot traffic patterns which concentrate the packets into a particular section of the network will cause imbalances in the amount of work handled by each thread.
-Even with the shared channel semaphore and the removal of the cycle barrier, the lightly loaded threads will almost always be stalled waiting for the hotspot threads.
+Even with the shared channel semaphore and the removal of the cycle barrier, the lightly loaded threads will  always be stalled waiting for the hotspot threads.
 Fortunately, common network traffic patterns are balanced and network routing algorithm are usually designed to reduce hotspots and bottlenecks in the network.

### 2009-03-16 doc/paper.tex
I am good with it
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1158 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 only adds to the computational bottleneck and renders them unable to take
-advantage of new multicore processors.  This paper demonstrates how
+advantage of the new multicore processors.  This paper demonstrates how
 an interconnect network simulator can be parallelized into a multithreaded
@@
 development time and better designs as architects can sweep additional
-parameter spaces and consider more exotic configurations.}
+parameter spaces and consider a wider range of network configurations.}
 %}
@@
 In addition to the need for parallel programs, another major ramification of the multicore revolution has been the
-increased importance of interconnect network design.  As the number of
+increased importance of interconnection network design.  As the number of
 cores being placed on a chip increases, the processor interconnect plays a bigger and
@@
 The traditional bus architecture is no longer the appropriate medium of communication for large number of cores.
-Even more recently, single crossbar communication hubs will also not be sufficient in the future since the complexity of a single crossbar grows quadratically with the number of cores.
+Even more recently, single crossbar communication hubs will also not be sufficient in the future since the cost of a single crossbar grows quadratically with the number of cores.
 Interconnection network can provide the scalable and efficient communication
@@
 items in the queue can only be written to atomically.  A diagram of how
-this works can be seen in Figure \ref{lfqueue}.
+this works can be seen in Figure \ref{lfqueue}.\\
+~\\
 \begin{figure}[h]
@@
 \end{figure}
+
 We initially implemented this AtomicReference object with a PThread
@@
 a pop operation occurred to the queue, a down operation would have to be
-performed first.  We also modified the routers such that for every cycle
+performed first. The value of the semaphore is initialized to latency of the channel. We also modified the routers such that for every cycle
 of simulation, they would perform at least one up operation on every
@@
 The last operation that we performed on our code was to fuse the inner loops in figure~\ref{fig:runtime} together.  The initial simulator implementation was written to make it
-easy to understand the code and not for performance.  We moved all these operations into a single
-'for' loop in order to help improve some of the temporal locality of the
+easy to understand the code and not for performance.  We moved all these operations into a single loop in order to help improve some of the temporal locality of the
 code.  Not surprisingly this operation allowed the system to take
@@
 \section{Results \label{results}}
-Having completed a parallel implementation of our network simulator, our
-next step was to analyze the performance of the simulator to determine how
+Having completed a parallel implementation of our network simulator, we first tested its functionality.
+We generated a series of latency-throughput plot for a dragonfly network running with different traffic patterns.
+The resulting plot was nearly identical to those of the sequential simulator.
+On average there was about less than 0.1\% difference between the two versions.
+The slight different was caused by the randomness of packet injection.
+\\
+~\\
+ Our next step was to analyze the performance of the simulator to determine how
 much speedup it achieved.  We began first by analyzing the performance of
@@
 ~\\
+The first method of our program analysis was by using Intel's Vtune performance analyzer \cite{vtune}.
+Vtune provides a thread profiler for characterizing multi-threaded programs as well as access to all performance counters in Intel processors.
+The comparison between the profile of the parallel code and that of the sequential code did not reveal much insights.
+Majority of the execution time is still within the routing step during each simulation cycle.
+The lock free queues account for about 10\% of the runtime, proving that synchronization is a performance problem.
+One performance counter that did stand out is the data bus utilization.
+The utilization rate for the sequential code is 35\% while the parallel code with 2 threads was at 50\%.
+After tallying the number of cache lines transferred on the data bus (the BUS\_TRANS\_BURST.ALL\_AGENT counter) and use the formula provided by \cite{account}, we calculated that the parallel program uses approximated 3GB/s memory bandwidth from the front side bus.
+The maximum possible bandwidth on the system is 6.4GB/s, this was a further indication to us that simulator is perhaps memory bound.
+~\\
 To test our hypothesis that the simulator was memory bound, we ran the
@@
 the simulator remained poor and was even worse than the CMP version up to
-16 threads.  In addition to revealing that the poor performance was not
+16 threads.  In addition to revealing that the poor performance was perhaps not
 due to memory bandwidth, it also revealed that communication was playing a
@@
 ~\\
-After realizing that our program was not memory bound, we wanted to see if
+After putting the memory bandwidth hypothesis to doubt, we wanted to see if
 the size of the problem had any effect on the speedup.  We tested this by

### 2009-03-16 doc/sim.tex
I am good with it
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1158 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 The router objects contain all the data structure and operations necessary to route packets through the network.
-Similar to networks in real life, routers are mostly isolated from each other and can only communicate through channels or simple global signals.
+Similar to networks in real life, routers are mostly isolated from each other and can only communicate through the channels or simple global signals.
 As a result, the operations of individual routers are almost completely independent of other routers in the network.
@@
 However, the group's recent project involves simulating the dragonfly network \cite{cdragon}, which can easily grow to thousands or tens of thousands of nodes.
-Due to the organization of the simulator, the simulation runtime grows linearly with the number of network nodes and simulation time.
+Due to the organization of the simulator, the simulation runtime grows linearly with the number of network nodes and the number of simulated cycles.
 A typical simulation of the dragonfly network involves running a networks with 10K nodes for at least 10K cycles to capture the network's steady state.
@@
 Such long simulation time increase the turn around time for research and limits the number of experiments that can be run reasonably.
-Currently, the group utilizes program level parallelism by running multiple simulations simultaneously on the cluster.
+Currently, the group utilizes program level parallelism by running multiple simulations simultaneously on a cluster.
 But they are still limited by the runtime of each simulation.

### 2009-03-17 doc/disc.tex
Second revision of report.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1159 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 After operating on a component, the component is reinserted into the queue and the associated cycle is incremented.
-Under this organization, threads that have stalled ahead can pull work from other threads' work queues and help to finish the work in other network partitions.
+Under this organization, threads that have stalled ahead can pull work from other threads' work queues and help to finish the work in other network time slices.
 
 \subsection{Programming Model}
-Given the underwhelming speedup that we obtained through parallelization and the possibility of excessive false sharing in the program, this has lead us to think that maybe PThreads is not the most appropriate programming model for this simulator.
+Given the underwhelming speedup that we obtained through parallelization and the possibility of excessive false sharing in the program, this has lead us to think that PThreads is not the most appropriate programming model for this simulator.
 As we have stated through out the report, the only points of sharing between threads are the channels that bridge routers running under different threads.
@@
 But with the shared memory model of PThreads, false sharing due to the organization of data structure can easily occur.
-Since we know the exact points of data sharing and since the smart partitioning of the networks should keep the amount of inter-thread communication small, perhaps message passing programming models such as MPI will offer more scalability for the parallel simulator.
+Since we know the exact points of data sharing and since the smart partitioning of the networks should keep the amount of inter-thread communication small, perhaps message passing programming models such as MPI could offer more scalability for the parallel simulator.
 Shared channel objects can be written on top of MPI's send and receive interfaces, and many packets can be bundled together to reduce messaging overhead.

### 2009-03-17 doc/paper.tex
Second revision of report.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1159 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 to design and produce multicore chips that are capable of executing
-multithreaded programs.  While this has greatly increased the computation power of the processors, existing programs that have been
+multithreaded programs.  While this has greatly increased the computational power of the processors, existing programs that have been
 written in a single threaded programming model must be rewritten in a
@@
 of data.
-The traditional bus architecture is no longer the appropriate medium of communication for large number of cores.
+The traditional bus architecture is no longer the appropriate medium of communication for a large number of cores.
 Even more recently, single crossbar communication hubs will also not be sufficient in the future since the cost of a single crossbar grows quadratically with the number of cores.
-Interconnection network can provide the scalable and efficient communication
+Interconnection networks can provide the scalable and efficient communication
 infrastructures for future multicore processors. \\
@@
 In order to parallelize the network simulator, we first set about
-identifying sources of parallelism that we could exploit.  As mentioned previous, the routers in a network operate relatively autonomously.   Each router in the network can be simulated
+identifying sources of parallelism that we could exploit.  As mentioned previously, the routers in a network operate relatively autonomously.   Each router in the network can be simulated
 independently as long as the communication that occurs between the routers
@@
 to access the queue, it first had to acquire the lock and then release the
-lock when finished performing its operation on the queue.\\
+lock when it finished performing its operation on the queue.\\
 ~\\
@@
 However, this method adds an additional layer of complexity for the user.
-For each topology in the simulator, we need to write a network partitioning algorithm for that topology.
+For each topology in the simulator, we needed to write a network partitioning algorithm for that topology.
 Ideally we would like a generalized algorithm that would discover the optimal partitioning to minimize locking given any network, but this is not within the scope of this project and is a very difficult (NP-Complete) problem by itself.
@@
 Having completed a parallel implementation of our network simulator, we first tested its functionality.
-We generated a series of latency-throughput plot for a dragonfly network running with different traffic patterns.
+We generated a series of latency-throughput plots for a dragonfly network running with different traffic patterns.
 The resulting plot was nearly identical to those of the sequential simulator.
-On average there was about less than 0.1\% difference between the two versions.
+On average there was less than 0.1\% difference between the two versions.
 The slight different was caused by the randomness of packet injection.
@@
 all the queues were replaced with lock-free queues.  However, once we were
-able to only implement the shared channels with lock-free queues , we could see a noticeable performance gain and even some
+able to only implement the shared channels with lock-free queues, we could see a noticeable performance gain and even some
 additional scalability.  Finally, once we fused several of the loops being
@@
 set about attempting to determine the reason for such poor performance.
-Our initial hypothesis was that application was memory bound based on the
+Our initial hypothesis was that the application was memory bound based on the
 fact that there is a significant amount of data being processed in our
@@
 programming model implies that if the working set cannot fit inside the L2
-cache, then there will be a significant amount of data being moved both to
+cache, then there would be a significant amount of data being moved both to
 and from main memory.\\
 ~\\
-The first method of our program analysis was by using Intel's Vtune performance analyzer \cite{vtune}.
+The first method of our program analysis was to use Intel's Vtune performance analyzer \cite{vtune}.
 Vtune provides a thread profiler for characterizing multi-threaded programs as well as access to all performance counters in Intel processors.
 The comparison between the profile of the parallel code and that of the sequential code did not reveal much insights.
-Majority of the execution time is still within the routing step during each simulation cycle.
-The lock free queues account for about 10\% of the runtime, proving that synchronization is a performance problem.
+The majority of the execution time was still within the routing step during each simulation cycle.
+The lock free queues account for about 10\% of the runtime, proving that synchronization is still a performance problem.
 One performance counter that did stand out is the data bus utilization.
-The utilization rate for the sequential code is 35\% while the parallel code with 2 threads was at 50\%.
-After tallying the number of cache lines transferred on the data bus (the BUS\_TRANS\_BURST.ALL\_AGENT counter) and use the formula provided by \cite{account}, we calculated that the parallel program uses approximated 3GB/s memory bandwidth from the front side bus.
-The maximum possible bandwidth on the system is 6.4GB/s, this was a further indication to us that simulator is perhaps memory bound.
+The utilization rate for the sequential code was 35\% while the parallel code with 2 threads was at 50\%.
+After tallying the number of cache lines transferred on the data bus (the BUS\_TRANS\_BURST.ALL\_AGENT counter) and use the formula provided by \cite{account}, we calculated that the parallel program used approximated 3GB/s memory bandwidth from the front side bus.
+The maximum possible bandwidth on the system is 6.4GB/s, this was a further indication to us that simulator was perhaps memory bound.
+\\
 ~\\
@@
 \end{figure}
+~\\
 From the result of this experiment we were able to see that bandwidth was
@@
 extent.  The results from this experiment can be seen in figure
-\ref{niagara}.
+\ref{niagara}.\\
 \begin{figure}[h]
@@
 \end{figure}
+~\\
 We noticed that in this experiment, both the T2 and T2+ were able to
@@
 began to fall off.  We determined that once the threads' working set sizes
-began to interfere with each other in the shared L2 cache is when the
-performance begins to degrade and couldn't offer us much insight into  any potential false sharing that was occurring when run on other
+began to interfere with each other in the shared L2 cache was when the
+performance began to degrade and couldn't offer us much insight into any potential false sharing that was occurring when run on other
 hardware platforms.

### 2009-03-17 doc/sim.tex
Second revision of report.
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1159 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 The simulator allows for a wide range of user customizations when constructing a network.
-The simulator supports a variety of network topologies, router micro-architectures, routing algorithms, traffic pattern, and simulation modes.
+The simulator supports a variety of network topologies, router micro-architectures, routing algorithms, traffic patterns, and simulation modes.
 The priority of the simulator is  to guarantee the correctness and accuracy of simulation, and as a result the simulator is implemented in a cycle-accurate fashion.
@@
 \includegraphics[width=3in]{simulator.eps}
-\caption{Simulator organization: Network is composed of routers and channels and simulations are run through the traffic manager. }
+\caption{Simulator organization: the network is composed of routers and channels and simulations are run through the traffic manager. }
 \label{fig:simulator}
@@
 Each iteration of this loop represents a cycle in the network.
-Within each network cycle, several steps occurs to advance the state of the network.
+Within each network cycle, several steps occur to advance the state of the network.
 At each step, the simulator loops through all the routers in the network to perform a specific function.
-As mentioned previously the operation of routers are mostly independent of each other.
-Therefore, there are no dependencies between each iterations of these loops.
+As mentioned previously the operation of a router is mostly independent of other routers.
+Therefore, there are no dependencies between each iteration of these loops.
 Channels are the only source of sharing between the routers

### 2009-03-17 doc/paper.tex
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1160 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 \usepackage{algorithmic}
-\setlength{\textwidth}{6.5in}
-\setlength{\textheight}{9.5in}
-\setlength{\topmargin}{0in}
-\setlength{\oddsidemargin}{0in}
-\setlength{\evensidemargin}{0in}
-\setlength{\topmargin}{0in}
-\setlength{\headheight}{0in}
-\setlength{\headsep}{0in}
-\setlength{\topskip}{0.5in}
+\usepackage[top=1in, bottom=1in, left=0.75in, right=0.75in]{geometry}
+\setlength{\columnsep}{2pc}
+\usepackage[small]{caption}
 %% be sure to change the:
@@
 \centering
-\includegraphics[width=3in, height=2.25in]{lfqueue.eps}
+\includegraphics[width=3in]{lfqueue.eps}
 \caption{This figure demonstrates the layout of the objects associated with a lock-free queue of length 2 (the sentinel value does not count to towards the length of the list).
@@
 \centering
-\includegraphics[width=3in,height=3in]{synch.eps}
+\includegraphics[width=3in]{synch.eps}
 \caption{\label{synch} This figure shows the performance of the different
@@
 \centering
-\includegraphics[height=3in,width=3in]{smp.eps}
+\includegraphics[width=3in]{smp.eps}
 \caption{\label{smp}This figure shows the performance of the simulator on
@@
 \centering
-\includegraphics[height=3in,width=3in]{probsize.eps}
+\includegraphics[width=3in]{probsize.eps}
 \caption{\label{probsize}This figure shows the speedup versus network size of a
@@
 \centering
-\includegraphics[height=3in,width=3in]{niagara.eps}
+\includegraphics[width=3in]{niagara.eps}
 \caption{\label{niagara}This figure shows the speedup of the parallel simulator

### 2009-04-09 doc/manual.tex
add Id tags
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@848 ce7aba0d-9f81-40a6-a495-cecae84b0515
remove bogus svn:executable properties
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@938 ce7aba0d-9f81-40a6-a495-cecae84b0515
edited
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@1202 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
+% $Id$
+
+% Copyright (c) 2007, Trustees of Leland Stanford Junior University
+% All rights reserved.
+% 
+% Redistribution and use in source and binary forms, with or without 
+% modification, are permitted provided that the following conditions are met:
+% 
+% Redistributions of source code must retain the above copyright notice, this 
+% list of conditions and the following disclaimer.
+% Redistributions in binary form must reproduce the above copyright notice, 
+% this list of conditions and the following disclaimer in the documentation 
+% and/or other materials provided with the distribution.
+% Neither the name of the Stanford University nor the names of its contributors
+% may be used to endorse or promote products derived from this software without
+% specific prior written permission.
+% 
+% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" 
+% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE 
+% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE 
+% ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE 
+% LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR 
+% CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF 
+% SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS 
+% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN 
+% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) 
+% ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE 
+% POSSIBILITY OF SUCH DAMAGE.
+
 \documentclass[11pt]{article}
@@
 
-The latest version of the simulator is available from
-\texttt{http://cva.stanford.edu} as a compressed tar archive.  UNIX/Linux
-users can extract this archive using the tar utility
+The latest version of the simulator can be checked out from Booksim's sourceforge subversion (SVN) repository. Make sure SVN is installed on your machine and run 
 \begin{verbatim}
-
-  tar xvfz booksim-2.0.tar.gz  
+svn co https://booksim.svn.sourceforge.net/svnroot/booksim booksim 
 \end{verbatim}
 
-Windows users can use a compression program such as Winrar to extract
-the archive.
+The simulator files should now be in \texttt{booksim/trunk/src/}. The simulator itself is written in C++ and has been specifically
+tested with GNU's G++ compiler (version $\ge3$).
+The front end of the simulator uses LEX and YACC generated parser to process the simulator configuration file.
+But if you do not plan on making changes to the front end parser, the LEX and YACC are not needed.
+The
+\texttt{Makefile} should be edited so that the first lines give the
+paths to the tools. The default \texttt{Makefile} should work on the Stanford Leland machines. Type \texttt{make} to build the simulator.
 
-The simulator itself is written in C++ and has been specifically
-tested with GNU's G++ compiler (version $\ge3$).  In addition, both a
-LEX and YACC tool (also known as FLEX and BISON) are needed to create
-the configuration parser.  These are standard tools in any UNIX/Linux
-development environment.  It is suggested that Windows users download
-the CYGWIN versions (\texttt{http://www.cygwin.com}) of these UNIX
-development tools to simplify their compilation process.  The
-\texttt{Makefile} should be edited so that the first lines give the
-paths to the tools.  At Stanford, for example, the compiler, YACC, and
-LEX are stored in the \texttt{/usr/pubsw/bin} directory.  The
-\texttt{Makefile} reflects this:
-\begin{verbatim}
-CPP    = /usr/pubsw/bin/g++
-YACC   = /usr/pubsw/bin/bison -d
-LEX    = /usr/pubsw/bin/flex
-\end{verbatim}
-Then, the simulator can be compiled by running \texttt{make} in the
-directory that contains the \texttt{Makefile}.
-
+For Windows users: the above instructions has been tested to work with the newest version of Cygwin. 
 \subsection{Running a simulation}
@@
 \end{verbatim}
-The optional parameter \texttt{configfile} is a file that contains
+The parameter \texttt{configfile} is a file that contains
 configuration information for the simulator.  So, for example, to
@@
 network on uniform traffic, a configuration such as the one shown in
-Figure~\ref{fig:config_example} could be used. This particular configuration is stored in examples/torus88.
+Figure~\ref{fig:config_example} could be used. This particular configuration is stored in \texttt{examples/torus88}.
 
@@
 contains basic information about the routing algorithm, flow control,
-and traffic.  This simple example uses dimension-order routing and and four
+and traffic.  This simple example uses dimension-order routing and four
 virtual channels.  The \texttt{injection\_rate} parameter
-is added to tell the simulator to inject 0.15 packets per simulation
+is added to tell the simulator to inject (on average) 0.15 packets per simulation
 cycle per node.  The packets are default to size of 1 flit.  Also, any line of the configuration
@@
 Section~\ref{sec:config_params}.
+Any parameters not specified by the user will take on a default value.
+The default value for every parameter in the simulator is in the file \texttt{booksim\_config.cpp}.
 
@@
 \begin{verbatim}
-% Average latency = 6.02008
-% Accepted packets = 0.11 at node 52 (avg = 0.147094)
-% latency change    = 1
-% throughput change = 1
->>>>>>> variant B
 BEGIN Configuration File
@@
 END Configuration File
-Unable to open flit watch file, continuing with simulation
 0%=================================
-% Average latency = 10.6252
-% Accepted packets = 0.119 at node 5 (avg = 0.148875)
-lat(1) = 10.6252;
+% Average latency = 10.3206
+% Accepted packets = 0.113 at node 38 (avg = 0.149031)
+lat(1) = 10.3206;
 bins = [ 0 ...999];
-freq = [ 0 0 124 16 497 71 920 166 1336 303 1553 436 1351 
-406 926 328 537 213 173 77 43 24 13 6 5 2 1 1 0 ...]
+freq = [ 0 0 145 4 550 52 988 137 1439 285 1656 428 1359 417 958 294 453 163 143
+ 49 10 7 1 0 ....
 
-...
+....
 
@@
 
-...
+....
 
@@
 % Draining remaining packets ...
-Time taken is 6031 cycles
+Time taken is 6030 cycles
 ====== Traffic class 0 ======
-Overall average latency = 10.658 (1 samples)
-Overall average accepted rate = 0.148383 (1 samples)
-Overall min accepted rate = 0.134279 (1 samples)
-Average hops = 5.01172 (57148 samples)
+Overall average latency = 10.2897 (1 samples)
+Overall average accepted rate = 0.14834 (1 samples)
+Overall min accepted rate = 0.129043 (1 samples)
+Average hops = 4.99839 (57153 samples)
 \end{verbatim}
@@
 
-\section{Examples}
+\section{Example}
 \label{sec:examples}
@@
 making this measurement in a 8-ary 2-mesh network under the transpose
-traffic pattern.  This configuration was used to generate Figure 25.2
+traffic pattern.  This configuration was used to generate a single data point in Figure 25.2
 in PPIN.  The particular configuration accounts for some small delays
@@
 sim_type       = latency;
-injection_rate = 0.005;
+injection_rate = 0.01;
 \end{verbatim}
@@
 configuration parameters --- a user can incorporate additional options
-by changing the \texttt{booksim\_config.cpp} file. Most configuration parameters have default values that should work without user specification
+by changing the \texttt{booksim\_config.cpp} file.
 
@@
 
-\item[flattened butterfly] A topology based on ***
+\item[flattened butterfly] A topology based on the paper "Flattened butterfly: a cost-efficient topology for high-radix networks" ISCA 2007
 
@@
 
-\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung.
+\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung from  EE382C several years ago.
 
@@
 \appendix
-
 \section{Random number generation}

### 2009-06-24 doc/manual.tex
update quick start section
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@1359 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 
-The latest version of the simulator can be checked out from Booksim's sourceforge subversion (SVN) repository. Make sure SVN is installed on your machine and run 
+The latest official release of the simulator can be checked out from Booksim's sourceforge subversion (SVN) repository. Make sure SVN is installed on your machine and run the following command:
 \begin{verbatim}
-svn co https://booksim.svn.sourceforge.net/svnroot/booksim booksim 
+svn co https://booksim.svn.sourceforge.net/svnroot/booksim
 \end{verbatim}
 
-The simulator files should now be in \texttt{booksim/trunk/src/}. The simulator itself is written in C++ and has been specifically
-tested with GNU's G++ compiler (version $\ge3$).
-The front end of the simulator uses LEX and YACC generated parser to process the simulator configuration file.
-But if you do not plan on making changes to the front end parser, the LEX and YACC are not needed.
-The
-\texttt{Makefile} should be edited so that the first lines give the
-paths to the tools. The default \texttt{Makefile} should work on the Stanford Leland machines. Type \texttt{make} to build the simulator.
+The simulator files should now be in the \texttt{booksim/trunk/src/} directory.
+The simulator itself is written in C++ and has been specifically tested with GNU's G++ compiler (version $\ge3$).
+The front end of the simulator uses LEX and YACC generated parser to process the simulator configuration file; however, unless you plan on making changes to the front end parser, LEX and YACC are not needed.
+The \texttt{Makefile} should be edited so that the first few lines reflect the correct paths to the tools for your particular system.
+The default \texttt{Makefile} should work on the Stanford Leland machines.
+Type \texttt{make} to build the simulator.
 
-For Windows users: the above instructions has been tested to work with the newest version of Cygwin. 
+A note for Windows users:
+The above instructions have been tested to work with the newest version of Cygwin.
+
 \subsection{Running a simulation}
@@
 
-The syntax of the simulator is simply
+The simulator is invoked using the following command line:
 \begin{verbatim}
@@
 \end{verbatim}
-The parameter \texttt{configfile} is a file that contains
-configuration information for the simulator.  So, for example, to
-simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube)
-network on uniform traffic, a configuration such as the one shown in
-Figure~\ref{fig:config_example} could be used. This particular configuration is stored in \texttt{examples/torus88}.
+
+The parameter \texttt{configfile} is a file that contains configuration information for the simulator.
+So, for example, to simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube) network on uniform traffic, a configuration such as the one shown in Figure~\ref{fig:config_example} could be used.
+This particular example configuration file can be found in the \texttt{examples/torus88} directory.
 
@@
 
-In addition to specifying the topology, the configuration file also
-contains basic information about the routing algorithm, flow control,
-and traffic.  This simple example uses dimension-order routing and four
-virtual channels.  The \texttt{injection\_rate} parameter
-is added to tell the simulator to inject (on average) 0.15 packets per simulation
-cycle per node.  The packets are default to size of 1 flit.  Also, any line of the configuration
-that begins with \texttt{//} is treated as a comment and ignored by
-the simulator.  A detailed list of configuration parameters is given in
-Section~\ref{sec:config_params}.
-Any parameters not specified by the user will take on a default value.
-The default value for every parameter in the simulator is in the file \texttt{booksim\_config.cpp}.
+In addition to specifying the topology, the configuration file also contains basic information about the routing algorithm, flow control, and traffic.
+This simple example uses dimension-order routing and four virtual channels.
+The \texttt{injection\_rate} parameter is added to tell the simulator to inject (on average) 0.15 packets per simulation cycle per node.
+Packet size defaults to a single flit.
+Also, any line of the configuration file that begins with \texttt{//} is treated as a comment and ignored by the simulator.
+A detailed list of configuration parameters is given in Section~\ref{sec:config_params}.
+Any parameters not specified by the user will take on default values.
+The default value for every parameter in the simulator is specified in the file \texttt{booksim\_config.cpp}.
 

### 2009-08-14 doc/manual.tex
update booksim manual
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@1427 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 
-\author{Ted Jiang, George Michelogiannakis, Daniel Becker, Brian Towles and William J. Dally}
+\author{Nan Jiang, George Michelogiannakis, Daniel Becker, Brian Towles and William J. Dally}
 \title{\simname 2.0 User's Guide}
@@
 
-\subsubsection[Batch mode]
+\subsubsection{Batch mode}
 Batch mode is enabled by setting \texttt{use\_read\_write= 1}.
@@
 
-\subsubsection[Traffic Patterns]
+\subsubsection{Traffic Patterns}
 The simulator also supports several different traffic patterns that

### 2010-03-09 doc/manual.tex
update license text
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@1535 ce7aba0d-9f81-40a6-a495-cecae84b0515
update book sim manual
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@1767 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 
-% Copyright (c) 2007, Trustees of Leland Stanford Junior University
+% Copyright (c) 2007-2009, Trustees of The Leland Stanford Junior University
 % All rights reserved.
@@
 
-The two traffic modes, injection rate or batch, can be set by the integer paramter \texttt[use\_read\_write], where 0 means injection mode. Injection mode is on by default.
-
 \subsubsection{Injection mode}
-The rate at which packkets are injected into the simulator is set using
+The rate at which packets are injected into the simulator is set using
 the \texttt{injection\_rate} option.  The simulator's cycle time is a
 flit cycle, the time it takes a single flit to be injected at a
-source, and the injection rate is specified in packets per flit cycle.
+source, and the injection rate is specified in packets per flit cycle\footnote{The unit of \texttt{injection\_rate} can be changed to flits per cycle by setting \texttt{injection\_rate\_uses\_flits} to 1.}.
 For example, setting \texttt{injection\_rate = 0.25} means that each
-source injects a new packet one of every four simulator cycles.  The
-injection process can also be specified as either Bernoulli
+source injects a new packet in one out of every four simulator cycles.
+
+The injection process can further be specified as either Bernoulli process 
 (\texttt{injection\_process = bernoulli}) or an on-off process
 (\texttt{injection\_process = on\_off}).  The burstiness of the latter
-injection process is controlled via the \texttt{burst\_alpha} and
-\texttt{burst\_beta} parameter.  See PPIN Section 24.2.2 for a
+is controlled via the \texttt{burst\_alpha} and
+\texttt{burst\_beta} parameters.  See PPIN Section 24.2.2 for a
 description of the on-off process and its parameters.
 
-The unit of injection is packets, which may be comprised of many
-
-flits. All packets in injection mode are identical. The number of flits per packet is set using the
+\subsubsection{Request-reply traffic}
+By default, all packets that are injected into the network have the same, 
+fixed length. The number of flits per packet is set using the
 \texttt{const\_flits\_per\_packet} option.
 
-\subsubsection{Batch mode}
-Batch mode is enabled by setting \texttt{use\_read\_write= 1}.
-In batch mode, \texttt[injection\_rate] paramter is ignored.
-Instead, traffic are injected into the network as fast as possible until \texttt{batch\_size} of request packets are generated by each node.
-When a destination node receives a request packet, it returns a reply packet.
-The total number of packets injected in batch mode is 2$\times$ \texttt{batch\_size} per node.
-The parameter \texttt[max\_outstanding] specifies the maximum number of (request or reply) packets waiting at a node to be injected into the network.
+Alternatively, the simulator can be configured to generate request-reply traffic.
+In this mode, the traffic manager injects read and write requests into the network; when a destination node receives such a request packet, it returns a reply packet of the same type\footnote{Injection of reply packets has priority over injection of new requests packets.}.
+Consequently, the effective overall network load is generated by both the injected requests and the automatically generated replies.
+Request-reply traffic ignores the \texttt{const\_flits\_per\_packet} option; instead, packet sizes are determined by the \texttt{\{read|write\}\_\{request|reply\}\_size} options.
+Furthermore, the mapping of packet types to VCs can be customized using the \texttt{\{read|write\}\_\{request|reply\}\_\{begin|end\}\_vc} options.
 
-\subsubsection{Traffic Patterns}
+\subsubsection{Traffic patterns}
 The simulator also supports several different traffic patterns that
@@
 the \texttt{perm\_seed} option.  So, randomly selecting values for
-\texttt{perm\_seed} gives a random sampling of permutation while a
+\texttt{perm\_seed} gives a random sampling of permutations while a
 fixed value of \texttt{perm\_seed} allows the same permutation to be

### 2010-03-09 doc/manual.tex
update book sim manual
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@1768 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@
 flit cycle, the time it takes a single flit to be injected at a
-source, and the injection rate is specified in packets per flit cycle\footnote{The unit of \texttt{injection\_rate} can be changed to flits per cycle by setting \texttt{injection\_rate\_uses\_flits} to 1.}.
+source, and the injection rate is specified in packets per flit cycle.
 For example, setting \texttt{injection\_rate = 0.25} means that each
 source injects a new packet in one out of every four simulator cycles.
+The unit of \texttt{injection\_rate} can optionally be changed to flits per cycle by 
+setting \texttt{injection\_rate\_uses\_flits} to 1.
 
@@
 
-Alternatively, the simulator can be configured to generate request-reply traffic.
-In this mode, the traffic manager injects read and write requests into the network; when a destination node receives such a request packet, it returns a reply packet of the same type\footnote{Injection of reply packets has priority over injection of new requests packets.}.
-Consequently, the effective overall network load is generated by both the injected requests and the automatically generated replies.
-Request-reply traffic ignores the \texttt{const\_flits\_per\_packet} option; instead, packet sizes are determined by the \texttt{\{read|write\}\_\{request|reply\}\_size} options.
-Furthermore, the mapping of packet types to VCs can be customized using the \texttt{\{read|write\}\_\{request|reply\}\_\{begin|end\}\_vc} options.
+Alternatively, the simulator can be configured to generate request-reply 
+traffic. In this mode, the traffic manager injects read and write requests 
+into the network; when a destination node receives such a request 
+packet, it returns a reply packet of the same type. Injection of reply 
+packets has priority over injection of new requests packets. 
+Consequently, the effective overall network load is generated by both 
+the injected requests and the automatically generated replies. 
+Request-reply traffic ignores the \texttt{const\_flits\_per\_packet} 
+option; instead, packet sizes are determined by the 
+\texttt{\{read|write\}\_\{request|reply\}\_size} options. Furthermore, 
+the mapping of packet types to VCs can be customized using the 
+\texttt{\{read|write\}\_\{request|reply\}\_\{begin|end\}\_vc} options.
 

### 2011-04-26 src/allocators/islip.C
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/gems_interface@3426 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,7 +1,7 @@
 // $Id: islip.cpp 1839 2010-03-24 02:03:56Z dub $
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/networks/network.C

@@ -1,7 +1,7 @@
 // $Id: network.cpp 1970 2010-05-10 11:50:15Z dub $
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/networks/singlenet.C

@@ -1,7 +1,7 @@
 // $Id: singlenet.cpp 1970 2010-05-10 11:50:15Z dub $
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/vc.C

@@ -1,7 +1,7 @@
 // $Id: vc.cpp 2195 2010-06-30 20:42:08Z qtedq $
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/allocators/wavefront.C
fix messed up copyright notices
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/gems_interface@3427 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,7 +1,7 @@
 // $Id: wavefront.cpp 1839 2010-03-24 02:03:56Z dub $
 
 /*
-Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/booksim_config.C

@@ -1,7 +1,7 @@
 // $Id: booksim_config.cpp 2520 2010-09-13 19:57:17Z dub $
 
 /*
-Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/networks/singlenet.C

@@ -1,7 +1,7 @@
 // $Id: singlenet.cpp 1970 2010-05-10 11:50:15Z dub $
 
 /*
-Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/outputset.C

@@ -1,7 +1,7 @@
 // $Id: outputset.cpp 2200 2010-07-01 19:33:38Z qtedq $
 
 /*
-Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/rng_wrapper.C

@@ -1,7 +1,7 @@
 // $Id: rng_wrapper.cpp 1839 2010-03-24 02:03:56Z dub $
 
 /*
-Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-04-26 src/vc.C

@@ -1,7 +1,7 @@
 // $Id: vc.cpp 2195 2010-06-30 20:42:08Z qtedq $
 
 /*
-Copyright (c) 2007-20[01][0-9], Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2011-05-16 doc/manual.tex
remove stray reference to 'priority' from BookSim 2.0 manual
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@1864 ce7aba0d-9f81-40a6-a495-cecae84b0515
remove reference to MECS topology (no longer supported)
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@2911 ce7aba0d-9f81-40a6-a495-cecae84b0515
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/classes@3426 ce7aba0d-9f81-40a6-a495-cecae84b0515
fix messed up copyright notices
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/cla

@@
 
-% Copyright (c) 2007-2009, Trustees of The Leland Stanford Junior University
+% Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 % All rights reserved.
@@
 
-The latest official release of the simulator can be checked out from Booksim's sourceforge subversion (SVN) repository. Make sure SVN is installed on your machine and run the following command:
+The latest official release of the simulator can be checked out from our subversion (SVN) repository.
+Make sure a subversion client is installed on your machine; under UNIX, you can use the following command to check out a working copy:
+
 \begin{verbatim}
-svn co https://booksim.svn.sourceforge.net/svnroot/booksim
+svn co https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0
 \end{verbatim}
 
-The simulator files should now be in the \texttt{booksim/trunk/src/} directory.
-The simulator itself is written in C++ and has been specifically tested with GNU's G++ compiler (version $\ge3$).
+The simulator files should now be in the \texttt{booksim2.0/trunk/src/} directory.
+The simulator itself is written in C++ and has been specifically tested with the GNU G++ compiler (version $\ge3$).
 The front end of the simulator uses LEX and YACC generated parser to process the simulator configuration file; however, unless you plan on making changes to the front end parser, LEX and YACC are not needed.
@@
 
-\item[MECS] A topology based on ***. Currently it only supports single-flit packets.
-
 \end{opt_list}
@@
 \item[select] Priority-based allocator.  Allocation is performed as in
-iSLIP, but with preference towards higher priority packets (see
-\texttt{priority} option in Section~\ref{sec:traffic}).
+iSLIP, but with preference towards higher priority packets.
+% (see \texttt{priority} option in Section~\ref{sec:traffic}).
 

### 2011-07-08 src/networks/flatfly_onchip.hpp
adaptiev routing stuff
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/indirect_adaptive_routing@3808 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,7 +1,7 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,
@@ -67,7 +67,8 @@ public:
   double Capacity( ) const;
   void InsertRandomFaults( const Configuration &config );
 };
-
+void adaptive_xyyx_flatfly( const Router *r, const Flit *f, int in_channel, 
+		  OutputSet *outputs, bool inject );
 void xyyx_flatfly( const Router *r, const Flit *f, int in_channel, 
 		  OutputSet *outputs, bool inject );
 void min_flatfly( const Router *r, const Flit *f, int in_channel, 

### 2011-07-08 src/routefunc.cpp

@@ -1,7 +1,7 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,
@@ -71,42 +71,6 @@ int gReadReqBeginVC, gReadReqEndVC;
 int gWriteReqBeginVC, gWriteReqEndVC;
 int gReadReplyBeginVC, gReadReplyEndVC;
 int gWriteReplyBeginVC, gWriteReplyEndVC;
-
-// ----------------------------------------------------------------------
-//
-//   Crossbar Network 
-//
-// ----------------------------------------------------------------------
-void dest_tag_crossbar( const Router *r, 
-			const Flit *f, 
-			int in_channel, 
-			OutputSet *outputs, 
-			bool inject ) {
-  
-  outputs->Clear() ;
-
-  // Output port determined by those bits of destination above 
-  //  concentration bits
-  int out_port = inject ? 0 : ( f->dest >> log_two( gC ) ) ;
-
-  int vcBegin = 0, vcEnd = gNumVCs-1;
-  if ( f->type == Flit::READ_REQUEST ) {
-    vcBegin = gReadReqBeginVC;
-    vcEnd = gReadReqEndVC;
-  } else if ( f->type == Flit::WRITE_REQUEST ) {
-    vcBegin = gWriteReqBeginVC;
-    vcEnd = gWriteReqEndVC;
-  } else if ( f->type ==  Flit::READ_REPLY ) {
-    vcBegin = gReadReplyBeginVC;
-    vcEnd = gReadReplyEndVC;
-  } else if ( f->type ==  Flit::WRITE_REPLY ) {
-    vcBegin = gWriteReplyBeginVC;
-    vcEnd = gWriteReplyEndVC;
-  }
-  assert(((f->vc >= vcBegin) && (f->vc <= vcEnd)) || (inject && (f->vc < 0)));
-
-  outputs->AddRange( out_port, vcBegin, vcEnd );
-}
 
 // ============================================================
 //  QTree: Nearest Common Ancestor
@@ -323,58 +287,34 @@ void fattree_nca( const Router *r, const Flit *f,
   } else {
     
     int dest = f->dest;
-    int router_id = r->GetID();
-    int router_depth, router_port;
-
-    int routers_so_far = 0, routers;
[…]

### 2011-07-08 src/routers/iq_router.hpp

@@ -1,7 +1,7 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,
@@ -56,8 +56,11 @@ class IQRouter : public Router {
   int _vcs;
   int _classes;
 
+  bool _vc_busy_when_full;
+
   bool _speculative;
   bool _spec_check_elig;
+  bool _spec_check_cred;
   bool _spec_mask_by_reqs;
   
   bool _active;
@@ -91,6 +94,7 @@ class IQRouter : public Router {
 
   tRoutingFunction   _rf;
 
+  int _output_buffer_size;
   vector<queue<Flit *> > _output_buffer;
 
   vector<queue<Credit *> > _credit_buffer;
@@ -148,7 +152,7 @@ public:
   
   void Display( ostream & os = cout ) const;
 
-  virtual int GetCredit(int out, int vc_begin, int vc_end ) const;
+  virtual int GetUsedCredit(int out, int vc_begin = -1, int vc_end = -1 ) const;
   virtual int GetBuffer(int i = -1) const;
   virtual vector<int> GetBuffers(int i = -1) const;
 

### 2011-07-08 src/stats.cpp

@@ -1,7 +1,7 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,
@@ -47,10 +47,9 @@ SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 
 Stats::Stats( Module *parent, const string &name,
 	      double bin_size, int num_bins ) :
-  Module( parent, name ),
-   _num_samples(0), _sample_sum(0.0), _sample_squared_sum(0.0), _min(numeric_limits<double>::max()), _max(-numeric_limits<double>::max()), _num_bins( num_bins ), _bin_size( bin_size ), _hist(_num_bins, 0)
+  Module( parent, name ), _num_bins( num_bins ), _bin_size( bin_size )
 {
-
+  Clear();
 }
 
 void Stats::Clear( )
@@ -61,8 +60,8 @@ void Stats::Clear( )
 
   _hist.assign(_num_bins, 0);
 
-  _min = numeric_limits<double>::max();
-  _max = -numeric_limits<double>::max();
+  _min = numeric_limits<double>::quiet_NaN();
+  _max = -numeric_limits<double>::quiet_NaN();
   
   //  _reset = true;
 }
@@ -104,29 +103,18 @@ int Stats::NumSamples( ) const
 
 void Stats::AddSample( double val )
 {
-  int b;
-
-  _num_samples++;
+  ++_num_samples;
   _sample_sum += val;
 
-  if(_num_samples == 0) {
-    _min = val;
-    _max = val;
-  } else {
-    if(val > _max) _max = val;
-    if(val < _min) _min = val;
-  }
+  // NOTE: the negation ensures that NaN values are handled correctly!
+  _max = !(val <= _max) ? val : _max;
+  _min = !(val >= _min) ? val : _min;
 
   //double clamp between 0 and num_bins-1
-  b = (int)fmax(floor( val / _bin_size ), 0.0);
+  int b = (int)fmax(floor( val / _bin_size ), 0.0);
   b = (b >= _num_bins) ? (_num_bins - 1) : b;
 
   _hist[b]++;
-}
-
-void Stats::AddSample( int val )
-{
-  AddSample( (double)val );
 }
 
[…]

### 2011-07-08 src/vc.cpp

@@ -1,7 +1,7 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2010, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,
@@ -43,27 +43,17 @@ SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 #include "booksim.hpp"
 #include "vc.hpp"
 
-int VC::total_cycles = 0;
 const char * const VC::VCSTATE[] = {"idle",
 				    "routing",
 				    "vc_alloc",
 				    "active"};
 
-VC::state_info_t VC::state_info[] = {{0},
-				     {0},
-				     {0},
-				     {0}};
-int VC::occupancy = 0;
-
 VC::VC( const Configuration& config, int outputs, 
 	Module *parent, const string& name )
   : Module( parent, name ), 
-    _state(idle), _state_time(0), _out_port(-1), _out_vc(-1), _total_cycles(0),
-    _vc_alloc_cycles(0), _active_cycles(0), _idle_cycles(0), _routing_cycles(0),
-    _pri(0), _watched(false), _expected_pid(-1), _last_id(-1), _last_pid(-1)
+    _state(idle), _out_port(-1), _out_vc(-1), _pri(0), _watched(false), 
+    _expected_pid(-1), _last_id(-1), _last_pid(-1)
 {
-  _size = config.GetInt( "vc_buf_size" ) + config.GetInt( "shared_buf_size" );
-
   _route_set = new OutputSet( );
 
   string priority = config.GetStr( "priority" );
@@ -87,7 +77,7 @@ VC::~VC()
   delete _route_set;
 }
 
-bool VC::AddFlit( Flit *f )
+void VC::AddFlit( Flit *f )
 {
   assert(f);
 
@@ -97,7 +87,6 @@ bool VC::AddFlit( Flit *f )
       err << "Received flit " << f->id << " with unexpected packet ID: " << f->pid 
 	  << " (expected: " << _expected_pid << ")";
       Error(err.str());
-      return false;
     } else if(f->tail) {
       _expected_pid = -1;
     }
@@ -105,11 +94,6 @@ bool VC::AddFlit( Flit *f )
     _expected_pid = f->pid;
   }
     
-  if((int)_buffer.size() >= _size) {
-    Error("Flit buffer overflow.");
-    return false;
[…]

### 2011-10-22 src/allocators/separable_input_first.cpp
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/classes@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/networks/anynet.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/booksim_config.cpp
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/systemsim_interface@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/traffic.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/allocators/pim.C
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/gems_interface@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,31 +1,28 @@
 // $Id: pim.cpp 1839 2010-03-24 02:03:56Z dub $
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
[…]

### 2011-10-22 src/credit.hpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/module.C

@@ -1,31 +1,28 @@
 // $Id: module.cpp 1839 2010-03-24 02:03:56Z dub $
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
[…]

### 2011-10-22 src/networks/qtree.C

@@ -1,31 +1,28 @@
 // $Id: qtree.cpp 2314 2010-07-25 02:06:38Z qtedq $
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
[…]

### 2011-10-22 src/routers/chaos_router.C

@@ -1,31 +1,28 @@
 // $Id: chaos_router.cpp 1969 2010-05-10 11:09:22Z dub $
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
[…]

### 2011-10-22 src/traffic.hpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/config_utils.cpp
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/indirect_adaptive_routing@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,30 +1,27 @@
 // $Id$
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/main.cpp
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/rng_wrapper.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 doc/manual.tex
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/gems_interface@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/pbooksim@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

--- 2010-03-09 doc/manual.tex
@@
 % and/or other materials provided with the distribution.
-% Neither the name of the Stanford University nor the names of its contributors
-% may be used to endorse or promote products derived from this software without
-% specific prior written permission.
 %
@@
 
-\author{Nan Jiang, George Michelogiannakis, Daniel Becker, Brian Towles and William J. Dally}
+\author{Ted Jiang, George Michelogiannakis, Daniel Becker, Brian Towles and William J. Dally}
 \title{\simname 2.0 User's Guide}
@@
 
-The latest official release of the simulator can be checked out from Booksim's sourceforge subversion (SVN) repository. Make sure SVN is installed on your machine and run the following command:
+The latest version of the simulator is available from
+\texttt{http://cva.stanford.edu} as a compressed tar archive.  UNIX/Linux
+users can extract this archive using the tar utility
 \begin{verbatim}
-svn co https://booksim.svn.sourceforge.net/svnroot/booksim
+
+  tar xvfz booksim-2.0.tar.gz  
 \end{verbatim}
 
-The simulator files should now be in the \texttt{booksim/trunk/src/} directory.
-The simulator itself is written in C++ and has been specifically tested with GNU's G++ compiler (version $\ge3$).
-The front end of the simulator uses LEX and YACC generated parser to process the simulator configuration file; however, unless you plan on making changes to the front end parser, LEX and YACC are not needed.
-The \texttt{Makefile} should be edited so that the first few lines reflect the correct paths to the tools for your particular system.
-The default \texttt{Makefile} should work on the Stanford Leland machines.
-Type \texttt{make} to build the simulator.
+Windows users can use a compression program such as Winrar to extract
+the archive.
 
-A note for Windows users:
-The above instructions have been tested to work with the newest version of Cygwin.
+The simulator itself is written in C++ and has been specifically
+tested with GNU's G++ compiler (version $\ge3$).  In addition, both a
+LEX and YACC tool (also known as FLEX and BISON) are needed to create
+the configuration parser.  These are standard tools in any UNIX/Linux
+development environment.  It is suggested that Windows users download
+the CYGWIN versions (\texttt{http://www.cygwin.com}) of these UNIX
+development tools to simplify their compilation process.  The
+\texttt{Makefile} should be edited so that the first lines give the
+paths to the tools.  At Stanford, for example, the compiler, YACC, and
+LEX are stored in the \texttt{/usr/pubsw/bin} directory.  The
+\texttt{Makefile} reflects this:
+\begin{verbatim}
+CPP    = /usr/pubsw/bin/g++
+YACC   = /usr/pubsw/bin/bison -d
+LEX    = /usr/pubsw/bin/flex
+\end{verbatim}
+Then, the simulator can be compiled by running \texttt{make} in the
+directory that contains the \texttt{Makefile}.
 
@@
 
-The simulator is invoked using the following command line:
+The syntax of the simulator is simply
 \begin{verbatim}
@@
 \end{verbatim}
-
-The parameter \texttt{configfile} is a file that contains configuration information for the simulator.
-So, for example, to simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube) network on uniform traffic, a configuration such as the one shown in Figure~\ref{fig:config_example} could be used.
-This particular example configuration file can be found in the \texttt{examples/torus88} directory.
+The optional parameter \texttt{configfile} is a file that contains
+configuration information for the simulator.  So, for example, to
+simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube)
+network on uniform traffic, a configuration such as the one shown in
+Figure~\ref{fig:config_example} could be used. This particular configuration is stored in examples/torus88.
 
@@
 
-In addition to specifying the topology, the configuration file also contains basic information about the routing algorithm, flow control, and traffic.
-This simple example uses dimension-order routing and four virtual channels.
-The \texttt{injection\_rate} parameter is added to tell the simulator to inject (on average) 0.15 packets per simulation cycle per node.
-Packet size defaults to a single flit.
-Also, any line of the configuration file that begins with \texttt{//} is treated as a comment and ignored by the simulator.
-A detailed list of configuration parameters is given in Section~\ref{sec:config_params}.
-Any parameters not specified by the user will take on default values.
-The default value for every parameter in the simulator is specified in the file \texttt{booksim\_config.cpp}.
+In addition to specifying the topology, the configuration file also
+contains basic information about the routing algorithm, flow control,
+and traffic.  This simple example uses dimension-order routing and and four
+virtual channels.  The \texttt{injection\_rate} parameter
+is added to tell the simulator to inject 0.15 packets per simulation
+cycle per node.  The packets are default to size of 1 flit.  Also, any line of the configuration
+that begins with \texttt{//} is treated as a comment and ignored by
+the simulator.  A detailed list of configuration parameters is given in
+Section~\ref{sec:config_params}.
 
@@
 \begin{verbatim}
+% Average latency = 6.02008
+% Accepted packets = 0.11 at node 52 (avg = 0.147094)
+% latency change    = 1
+% throughput change = 1
+>>>>>>> variant B
 BEGIN Configuration File
@@
 END Configuration File
+Unable to open flit watch file, continuing with simulation
 0%=================================
-% Average latency = 10.3206
-% Accepted packets = 0.113 at node 38 (avg = 0.149031)
-lat(1) = 10.3206;
+% Average latency = 10.6252
+% Accepted packets = 0.119 at node 5 (avg = 0.148875)
+lat(1) = 10.6252;
 bins = [ 0 ...999];
-freq = [ 0 0 145 4 550 52 988 137 1439 285 1656 428 1359 417 958 294 453 163 143
- 49 10 7 1 0 ....
+freq = [ 0 0 124 16 497 71 920 166 1336 303 1553 436 1351 
+406 926 328 537 213 173 77 43 24 13 6 5 2 1 1 0 ...]
 
-....
+...
 
@@
 
-....
+...
 
@@
 % Draining remaining packets ...
-Time taken is 6030 cycles
+Time taken is 6031 cycles
 ====== Traffic class 0 ======
-Overall average latency = 10.2897 (1 samples)
-Overall average accepted rate = 0.14834 (1 samples)
-Overall min accepted rate = 0.129043 (1 samples)
-Average hops = 4.99839 (57153 samples)
+Overall average latency = 10.658 (1 samples)
+Overall average accepted rate = 0.148383 (1 samples)
+Overall min accepted rate = 0.134279 (1 samples)
+Average hops = 5.01172 (57148 samples)
 \end{verbatim}
@@
 
-\section{Example}
+\section{Examples}
 \label{sec:examples}
@@
 making this measurement in a 8-ary 2-mesh network under the transpose
-traffic pattern.  This configuration was used to generate a single data point in Figure 25.2
+traffic pattern.  This configuration was used to generate Figure 25.2
 in PPIN.  The particular configuration accounts for some small delays
@@
 sim_type       = latency;
-injection_rate = 0.01;
+injection_rate = 0.005;
 \end{verbatim}
@@
 
-\item[flattened butterfly] A topology based on the paper "Flattened butterfly: a cost-efficient topology for high-radix networks" ISCA 2007
+\item[flattened butterfly] A topology based on ***
 
@@
 
-\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung from  EE382C several years ago.
+\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung.
 
@@
 
+The two traffic modes, injection rate or batch, can be set by the integer paramter \texttt[use\_read\_write], where 0 means injection mode. Injection mode is on by default.
+
 \subsubsection{Injection mode}
-The rate at which packets are injected into the simulator is set using
+The rate at which packkets are injected into the simulator is set using
 the \texttt{injection\_rate} option.  The simulator's cycle time is a
@@
 For example, setting \texttt{injection\_rate = 0.25} means that each
-source injects a new packet in one out of every four simulator cycles.
-The unit of \texttt{injection\_rate} can optionally be changed to flits per cycle by 
-setting \texttt{injection\_rate\_uses\_flits} to 1.
-
-The injection process can further be specified as either Bernoulli process 
+source injects a new packet one of every four simulator cycles.  The
+injection process can also be specified as either Bernoulli
 (\texttt{injection\_process = bernoulli}) or an on-off process
 (\texttt{injection\_process = on\_off}).  The burstiness of the latter
-is controlled via the \texttt{burst\_alpha} and
-\texttt{burst\_beta} parameters.  See PPIN Section 24.2.2 for a
+injection process is controlled via the \texttt{burst\_alpha} and
+\texttt{burst\_beta} parameter.  See PPIN Section 24.2.2 for a
 description of the on-off process and its parameters.
 
-\subsubsection{Request-reply traffic}
-By default, all packets that are injected into the network have the same, 
-fixed length. The number of flits per packet is set using the
+The unit of injection is packets, which may be comprised of many
+
+flits. All packets in injection mode are identical. The number of flits per packet is set using the
 \texttt{const\_flits\_per\_packet} option.
 
-Alternatively, the simulator can be configured to generate request-reply 
-traffic. In this mode, the traffic manager injects read and write requests 
-into the network; when a destination node receives such a request 
-packet, it returns a reply packet of the same type. Injection of reply 
-packets has priority over injection of new requests packets. 
-Consequently, the effective overall network load is generated by both 
-the injected requests and the automatically generated replies. 
-Request-reply traffic ignores the \texttt{const\_flits\_per\_packet} 
-option; instead, packet sizes are determined by the 
-\texttt{\{read|write\}\_\{request|reply\}\_size} options. Furthermore, 
-the mapping of packet types to VCs can be customized using the 
-\texttt{\{read|write\}\_\{request|reply\}\_\{begin|end\}\_vc} options.
+\subsubsection[Batch mode]
+Batch mode is enabled by setting \texttt{use\_read\_write= 1}.
+In batch mode, \texttt[injection\_rate] paramter is ignored.
+Instead, traffic are injected into the network as fast as possible until \texttt{batch\_size} of request packets are generated by each node.
+When a destination node receives a request packet, it returns a reply packet.
+The total number of packets injected in batch mode is 2$\times$ \texttt{batch\_size} per node.
+The parameter \texttt[max\_outstanding] specifies the maximum number of (request or reply) packets waiting at a node to be injected into the network.
 
-\subsubsection{Traffic patterns}
+\subsubsection[Traffic Patterns]
 The simulator also supports several different traffic patterns that
@@
 the \texttt{perm\_seed} option.  So, randomly selecting values for
-\texttt{perm\_seed} gives a random sampling of permutations while a
+\texttt{perm\_seed} gives a random sampling of permutation while a
 fixed value of \texttt{perm\_seed} allows the same permutation to be
@@
 \appendix
+
 \section{Random number generation}

### 2011-10-22 src/allocators/allocator.hpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/allocators/wavefront.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/creditchannel.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/globals.hpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/networks/qtree.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/networks/singlenet.hpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/power_module.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/rng.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/rng.hpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/routers/MECSChannels.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/routers/MECSForwarder.cpp

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/misc_utils.hpp
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/age_based@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,31 +1,28 @@
 // $Id$
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/allocators/separable_input_first.hpp
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/mpibooksim@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,31 +1,28 @@
 // $Id: separable_input_first.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/allocators/wavefront.hpp

@@ -1,31 +1,28 @@
 // $Id: wavefront.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
[…]

### 2011-10-22 src/networks/mecs.cpp

@@ -1,31 +1,28 @@
 // $Id: mecs.cpp 1117 2009-02-25 07:38:52Z qtedq $
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
[…]

### 2011-10-22 src/networks/qtree.hpp

@@ -1,31 +1,28 @@
 // $Id: qtree.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2011-10-22 src/rng.cpp

@@ -1,31 +1,28 @@
 // $Id: rng.cpp 938 2008-12-12 03:06:32Z dub $
 
 /*
-Copyright (c) 2007, Trustees of Leland Stanford Junior University
-All rights reserved.
+ Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ All rights reserved.
 
-Redistribution and use in source and binary forms, with or without modification,
-are permitted provided that the following conditions are met:
+ Redistribution and use in source and binary forms, with or without
+ modification, are permitted provided that the following conditions are met:
 
-Redistributions of source code must retain the above copyright notice, this list
-of conditions and the following disclaimer.
-Redistributions in binary form must reproduce the above copyright notice, this 
-list of conditions and the following disclaimer in the documentation and/or 
-other materials provided with the distribution.
-Neither the name of the Stanford University nor the names of its contributors 
-may be used to endorse or promote products derived from this software without 
-specific prior written permission.
+ Redistributions of source code must retain the above copyright notice, this 
+ list of conditions and the following disclaimer.
+ Redistributions in binary form must reproduce the above copyright notice, this
+ list of conditions and the following disclaimer in the documentation and/or
+ other materials provided with the distribution.
 
-THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
-ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED 
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE 
-DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR 
-ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES 
[…]

### 2012-08-30 src/allocators/islip.hpp
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/mpibooksim@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,7 +1,7 @@
 // $Id: islip.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/allocators/separable_output_first.cpp

@@ -1,7 +1,7 @@
 // $Id: separable_output_first.cpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/arbiters/arbiter.hpp

@@ -1,7 +1,7 @@
 // $Id: arbiter.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/arbiters/matrix_arb.hpp

@@ -1,7 +1,7 @@
 // $Id: matrix_arb.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/arbiters/roundrobin_arb.cpp

@@ -1,7 +1,7 @@
 // $Id: roundrobin_arb.cpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/arbiters/roundrobin_arb.hpp

@@ -1,7 +1,7 @@
 // $Id: roundrobin_arb.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/networks/isolated_mesh.hpp

@@ -1,7 +1,7 @@
 // $Id: isolated_mesh.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/networks/network.hpp

@@ -1,7 +1,7 @@
 // $Id: network.hpp 1102 2009-02-20 01:58:29Z qtedq $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/pipefifo.hpp

@@ -1,7 +1,7 @@
 // $Id: pipefifo.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/power_module.cpp

@@ -1,7 +1,7 @@
 // $Id: power_module.cpp 887 2008-12-04 23:02:18Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/power_module.hpp

@@ -1,7 +1,7 @@
 // $Id: power_module.hpp 887 2008-12-04 23:02:18Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/rng_double_wrapper.cpp

@@ -1,7 +1,7 @@
 // $Id: rng_double_wrapper.cpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/rng_wrapper.cpp

@@ -1,7 +1,7 @@
 // $Id: rng_wrapper.cpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/routers/event_router.cpp

@@ -1,7 +1,7 @@
 // $Id: event_router.cpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/routers/iq_router.hpp

@@ -1,7 +1,7 @@
 // $Id: iq_router.hpp 938 2008-12-12 03:06:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 allocators/maxsize.cpp
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/network_clusters@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,7 +1,7 @@
 // $Id: maxsize.cpp 3427 2011-04-26 00:31:25Z dub $
 
 /*
-  Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+  Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
   All rights reserved.
 
   Redistribution and use in source and binary forms, with or without 

### 2012-08-30 allocators/wavefront.hpp

@@ -1,7 +1,7 @@
 // $Id: wavefront.hpp 4080 2011-10-22 23:11:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 buffer.hpp

@@ -1,7 +1,7 @@
 // $Id: buffer.hpp 4296 2012-02-15 18:37:41Z qtedq $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 buffer_state.cpp

@@ -1,7 +1,7 @@
 // $Id: buffer_state.cpp 4188 2011-12-06 20:38:43Z qtedq $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 config_utils.hpp

@@ -1,7 +1,7 @@
 // $Id: config_utils.hpp 4080 2011-10-22 23:11:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 main.cpp

@@ -1,7 +1,7 @@
 // $Id: main.cpp 4080 2011-10-22 23:11:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 networks/kncube.cpp

@@ -1,7 +1,7 @@
 // $Id: kncube.cpp 4080 2011-10-22 23:11:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 packet_reply_info.cpp

@@ -1,7 +1,7 @@
 // $Id: packet_reply_info.cpp 4080 2011-10-22 23:11:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 power/power_module.cpp

@@ -1,7 +1,7 @@
 // $Id: power_module.cpp 4080 2011-10-22 23:11:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 power/switch_monitor.hpp

@@ -1,7 +1,7 @@
 // $Id: switch_monitor.hpp 4080 2011-10-22 23:11:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 random_utils.hpp

@@ -1,7 +1,7 @@
 // $Id: random_utils.hpp 4080 2011-10-22 23:11:32Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 routers/iq_router.cpp

@@ -1,7 +1,7 @@
 // $Id: iq_router.cpp 4888 2012-07-12 03:49:39Z qtedq $
 
 /*
-  Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+  Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
   All rights reserved.
 
   Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/arbiters/arbiter.C
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/gems_interface@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,7 +1,7 @@
 // $Id: arbiter.cpp 2187 2010-06-29 22:11:59Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/arbiters/roundrobin_arb.C

@@ -1,7 +1,7 @@
 // $Id: roundrobin_arb.cpp 2203 2010-07-02 00:04:02Z qtedq $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/booksim_main.C

@@ -1,7 +1,7 @@
 // $Id: main.cpp 2842 2010-11-12 03:13:28Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/config_utils.C

@@ -1,6 +1,6 @@
 // $Id: config_utils.cpp 2516 2010-09-01 23:09:06Z qtedq $
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/gemstrafficmanager.hpp

@@ -1,7 +1,7 @@
 // $Id: trafficmanager.hpp 2203 2010-07-02 00:04:02Z qtedq $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/module.C

@@ -1,7 +1,7 @@
 // $Id: module.cpp 1839 2010-03-24 02:03:56Z dub $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/networks/flatfly_onchip.C

@@ -1,7 +1,7 @@
 // $Id: flatfly_onchip.cpp 2382 2010-08-02 19:14:09Z qtedq $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/networks/qtree.C

@@ -1,7 +1,7 @@
 // $Id: qtree.cpp 2314 2010-07-25 02:06:38Z qtedq $
 
 /*
- Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+ Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
  All rights reserved.
 
  Redistribution and use in source and binary forms, with or without

### 2012-08-30 utils/sweep.sh
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/hotspot_reservation@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -2,7 +2,7 @@
 
 # $Id$
 
-# Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+# Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
 # All rights reserved.
 #
 # Redistribution and use in source and binary forms, with or without

### 2012-08-30 src/arbiters/probabilistic_arb.hpp
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/priority_tree@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@ -1,7 +1,7 @@
 // $Id: roundrobin_arb.hpp 3510 2011-05-09 22:07:44Z qtedq $
 
 /*
-Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
 All rights reserved.
 
 Redistribution and use in source and binary forms, with or without modification,

### 2013-11-26 doc/manual.tex
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/backoff_alloc@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/pbooksim@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/systemsim_interface@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/priority_tree@5188 ce7aba0d-9f81-40a6-a

--- 2011-05-16 doc/manual.tex
@@
 
-% Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
+% Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
 % All rights reserved.
@@
 A note for Windows users:
-The above instructions have been tested to work with the newest version of Cygwin.
+The above instructions have been tested to work with Cygwin 1.7.18.
 
@@
 \begin{verbatim}
-BEGIN Configuration File
-Name: examples/torus88
+BEGIN Configuration File: examples/torus88
+
+....
+
 // Topology
@@
 // Flow control
-num_vcs = 4;
+num_vcs = 2;
 // Traffic
@@
 
-END Configuration File
-0%=================================
-% Average latency = 10.3206
-% Accepted packets = 0.113 at node 38 (avg = 0.149031)
-lat(1) = 10.3206;
-bins = [ 0 ...999];
-freq = [ 0 0 145 4 550 52 988 137 1439 285 1656 428 1359 417 958 294 453 163 143
- 49 10 7 1 0 ....
+END Configuration File: examples/torus88
+Class 0:
+Packet latency average = 33.2807
+	minimum = 7
+	maximum = 73
 
@@
 
-% Warmed up ...Time used is 3000 cycles
+Warmed up ...Time used is 2000 cycles
 
@@
 
-% Draining all recorded packets ...
-% Draining remaining packets ...
-Time taken is 6030 cycles
+Draining all recorded packets ...
+Draining remaining packets ...
+Time taken is 5104 cycles
+====== Overall Traffic Statistics ======
 ====== Traffic class 0 ======
-Overall average latency = 10.2897 (1 samples)
-Overall average accepted rate = 0.14834 (1 samples)
-Overall min accepted rate = 0.129043 (1 samples)
-Average hops = 4.99839 (57153 samples)
+Packet latency average = 33.6964 (1 samples)
+	minimum = 7 (1 samples)
+	maximum = 84 (1 samples)
+
+....
+
+Total run time 1.01837
 \end{verbatim}
@@
 Then, the measurement phase begins and statistics continue to be
-reported after each sample period.  Once the measurement periods have
+reported after each sample period.  The measurement phase 
+typically last for last least 3 sample periods. 
+Once the measurement periods have
 passed, all the measurement packets are drained from the network
@@
 // Routing
-
-routing_function = dim_order;
+routing_function = dor;
 
 // Flow control
-
 num_vcs     = 8;
 vc_buf_size = 8;
-
 wait_for_tail_credit = 1;
@@
 // Router architecture
-
 vc_allocator = islip;
@@
 credit_delay   = 2;
-routing_delay  = 1;
+routing_delay  = 0;
 vc_alloc_delay = 1;
+sw_alloc_delay = 1;
 
@@
 // Traffic
-
 traffic = transpose;
-const_flits_per_packet = 20;
+packet_size = 20;
 
 // Simulation
+sim_type = latency;
 
-sim_type       = latency;
-injection_rate = 0.01;
+injection_rate = 0.005;
+
 \end{verbatim}
@@
 configuration file as illustrated by the example in 
-Section~\ref{sec:run_example}.  This section lists the existing
-configuration parameters --- a user can incorporate additional options
-by changing the \texttt{booksim\_config.cpp} file.
+Section~\ref{sec:run_example}.  This section lists the major
+configuration parameters for the simulator. Additional description can be found in the 
+configuration parameter class file \texttt{booksim\_config.cpp}.
+A user can incorporate additional options by changing the this file.
 
@@
 
-The \texttt{topology} parameter determines the underlying topology of the
-network. There is also a set of numerical parameters that describes the size of the networks.
+The \texttt{topology} parameter determines the underlying topology of
+the network. There is also a set of numerical parameters that
+describes the size of the networks.
 
@@
 \item[n] Network dimension
-\item[c] Network concentration, the number of nodes sharing a single router. Typically set to 1, >1 only in networks that has concentration (i.e. cmesh)
-\item[x] (NoC simulations only) The number of routers in the X dimension. Used to calculate channel latency between routers. 
-\item[y] (NoC simulations only) The number of routers in the y dimension. Used to calculate channel latency between routers. 
-\item[xr] (NoC simulations only) For networks that have c>1, the number of nodes in the x direction per router. Used to calculate channel latency between routers. 
-\item[yr] (NoC simulations only) For networks that have c>1, the number of nodes in the y direction per router. Used to calculate channel latency between routers. 
-\item[limit] Emulate partially filled networks.
-Nodes greater or equal to limit does not generate or receive packets during simulation.
-Though depending on the routing algorithm the routers associated with those nodes may be use to route passing by traffic.
+\item[c] Network concentration, the number of nodes sharing a single
+  router. Typically set to 1, >1 only in networks that has
+  concentration (i.e. cmesh).
+\item[x] (NoC simulations only) The number of routers in the X
+  dimension. Used to calculate channel latency between routers.  
+\item[y] (NoC simulations only) The number of routers in the y
+  dimension. Used to calculate channel latency between routers.  
+\item[xr] (NoC simulations only) For networks that have c>1, the
+  number of nodes in the x direction per router. Used to calculate
+  channel latency between routers.  
+\item[yr] (NoC simulations only) For networks that have c>1, the
+  number of nodes in the y direction per router. Used to calculate
+  channel latency between routers.
 
@@
 
-The channel latency of the network must be configured within the source code of the topology files. All topologies by default have channel latency of 1 cycles. Topologies available in \simname are:
+The channel latency of the network must be configured within the
+source code of the topology files. All topologies by default have
+channel latency of 1 cycles. Topologies available in \simname are:
 
@@
 parameter determines the network's radix and the \texttt{n} parameter
-determines the network's dimension.
+determines the network's dimension. Note: a $k$-ary 1-fly is
+essentially a single radix-k router, useful for testing.
 
@@
 
-\item[single] A network with a single node, used for testing single
-router performance.  The number of input and output ports for the node
-is determined by the \texttt{in\_ports} and \texttt{out\_ports} parameters,
-respectively.
-
 \item[torus] A $k$-ary $n$-cube (torus) topology.  The \texttt{k}
@@
 the network's dimension.
-\item[cmesh] Concentrated mesh topology is a A $k$-ary $n$-mesh topology with multiple nodes sharing a single router.
-The \texttt{c} determines the concentration.
-The cmesh topology has the option that turns on "express channels" as described in  by default these channels are turned off.
+\item[cmesh] Concentrated mesh topology is a A $k$-ary $n$-mesh
+  topology with multiple nodes sharing a single router. The \texttt{c}
+  determines the concentration. The cmesh topology has the option that
+  turns on "express channels" as described in  by default these
+  channels are turned off.
 
-\item[fat tree] Fat Tree topology with 3 levels.
-Nodes are routers are arranged in a tree structure but the number of links between levels stays constant.
-At the bottom \texttt{k} nodes shares a level 0 router.
+\item[fat tree] Fat Tree topology with 3 levels. Nodes are routers are
+  arranged in a tree structure but the number of links between levels
+  stays constant. At the bottom \texttt{k} nodes shares a level 0
+  router.
 
-\item[flattened butterfly] A topology based on the paper "Flattened butterfly: a cost-efficient topology for high-radix networks" ISCA 2007
+\item[flattened butterfly] A topology based on the paper "Flattened
+  butterfly: a cost-efficient topology for high-radix networks" ISCA
+  2007
+
+\item[dragonfly] A topology based on the paper ``Technology-driven,
+  highly-scalable dragonfly topology.'' ISCA 2008
 
@@
 
-\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung from  EE382C several years ago.
+\item[anynet] A topology based on an user input file specifying
+  connectivity of nodes and routers.
 
@@
 
-Both the \texttt{mesh} and \texttt{torus} topologies support the
-addition of random link failures with the \texttt{link\_failures}
-parameter.  The value of \texttt{link\_failures} determines the number
-of channels that are randomly removed from the topology and are thus
-no longer available for forwarding packets.  Moreover, the
-randomization for failed channels is controlled by selecting an
-integer value for the \texttt{fail\_seed} parameter --- a fixed seed
-gives a fixed set of failed channels, independent of other
-randomization in the simulation.  Also, note that only certain routing
-functions support this feature (see Section~\ref{sec:routing_algs}).
+%Both the \texttt{mesh} and \texttt{torus} topologies support the
+%addition of random link failures with the \texttt{link\_failures}
+%parameter.  The value of \texttt{link\_failures} determines the number
+%of channels that are randomly removed from the topology and are thus
+%no longer available for forwarding packets.  Moreover, the
+%randomization for failed channels is controlled by selecting an
+%integer value for the \texttt{fail\_seed} parameter --- a fixed seed
+%gives a fixed set of failed channels, independent of other
+%randomization in the simulation.  Also, note that only certain routing
+%functions support this feature (see Section~\ref{sec:routing_algs}).
 
@@
 
-The \texttt{physical\_subnetworks} parameter defines the number of physical sub-networks present in the network (defaults to one).
-All sub-networks receive the same configuration parameters and thus are identical. Traffic sources
-maintain an injection queue for each sub-network. The packet generation process is unaffected. It enqueues generated packets into the
-proper sub-network queue according to a division function in the traffic manager. At every cycle, flits at the head of each queue attempt to be injected.
-Traffic destinations can eject one flit from each sub-network each cycle.
+The \texttt{physical\_subnetworks} parameter defines the number of
+physical sub-networks present in the network (defaults to one). All
+sub-networks receive the same configuration parameters and thus are
+identical. Traffic sources maintain an injection queue for each
+sub-network. The packet 
+generation process is unaffected. It enqueues generated packets into
+the proper sub-network queue according to a division function in the
+traffic manager. At every cycle, flits at the head of each queue
+attempt to be injected. Traffic destinations can eject one flit from
+each sub-network each cycle.
 
@@
 for the topology.  Many routing algorithms need multiple virtual
-channels for deadlock freedom (VCDF).
-In addition to \texttt{routefunc.cpp}, some topologies source files include additional routing functions.
-Also, the simulator code is structured so that additional routing algorithms can be added with minimal changes to the overall simulator (see the \texttt{routefunc.cpp} file in the simulator's source code).
+channels for deadlock freedom. In addition to \texttt{routefunc.cpp},
+some topologies source files include additional routing
+functions. Also, the simulator code is structured so that additional
+routing algorithms can be added with minimal changes to the overall
+simulator (see the \texttt{routefunc.cpp} file in the simulator's
+source code).
 
@@
 
-\item[output\_delay] The processing delay incurred in the output queue
-of a router.
+%\item[output\_delay] The processing delay incurred in the output queue
+%of a router.
 \end{opt_list}
@@
 
-\item[speculative] Enable speculative switch allocation (i.e., allow switch allocation to occur in parallel with VC allocation for header flits).
+\item[speculative] Enable speculative switch allocation (i.e., allow
+  switch allocation to occur in parallel with VC allocation for header
+  flits).
 
-\item[filter\_spec\_grants] Determines how speculative grants are masked (\texttt{any\_nonspec\_gnts}: any non-speculative grant inhibits all speculative grants; \texttt{confl\_nonspec\_reqs}: speculative grants are inhibited by conflicting non-speculative requests; \texttt{confl\_nonspec\_gnts}: speculative grants are inhibited by conflicting non-speculative grants).
+%\item[filter\_spec\_grants] Determines how speculative grants are masked (\texttt{any\_nonspec\_gnts}: any non-speculative grant inhibits all speculative grants; \texttt{confl\_nonspec\_reqs}: speculative grants are inhibited by conflicting non-speculative requests; \texttt{confl\_nonspec\_gnts}: speculative grants are inhibited by conflicting non-speculative grants).
 
-\item[alloc\_iters] For the \texttt{islip}, \texttt{pim} and \texttt{select} allocators, allocation can be improved by performing multiple iterations of the algorithm; this parameter controls the number of iterations to be performed for both switch and VC allocation.
+\item[alloc\_iters] For the \texttt{islip}, \texttt{pim} and
+  \texttt{select} allocators, allocation can be improved by performing
+  multiple iterations of the algorithm; this parameter controls the
+  number of iterations to be performed for both switch and VC
+  allocation.
 
-\item[sw\_allocator] The type of allocator used for switch allocation.
-See Section~\ref{sec:alloc} for a list of the possible allocators.
+\item[arb\_type] If the VC or switch  allocator is a separable
+  input- or output-first allocator, this parameter selects the type of
+  arbiter to use.
 
-\item[sw\_alloc\_arb\_type] If the switch allocator is a separable input- or output-first allocator, this parameter selects the type of arbiter to use.
-
-\item[sw\_alloc\_iters] If the switch allocator is of type \texttt{islip}, \texttt{pim} or \texttt{select}, this parameter controls how many allocation iterations should be performed.
-This overrides the global \texttt{alloc\_iters} parameter.
+\item[sw\_allocator] The type of allocator used for switch
+  allocation. See Section~\ref{sec:alloc} for a list of the possible
+  allocators.
 
@@
 allocators.
-
-\item[vc\_alloc\_arb\_type] If the VC allocator is a separable input- or output-first allocator, this parameter selects the type of arbiter to use.
-
-\item[vc\_alloc\_iters] If the VC allocator is of type \texttt{islip}, \texttt{pim} or \texttt{select}, this parameter controls how many allocation iterations should be performed.
-This overrides the global \texttt{alloc\_iters} parameter.
 
@@
 
-\item[viewer\_trace] The simulator will generate very verbose print out of all activity inside the network.
+%\item[viewer\_trace] The simulator will generate very verbose print out of all activity inside the network.
 This print out should be fed into noc\_viewer for a graphic display of the activity inside the network.
